### 2021-12-26 paper.tex
init

\documentclass[12pt,titlepage,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{babel}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsthm}

\graphicspath{ {./images/} }
\DeclareMathOperator*{\E}{\mathbb{E}}

\begin{document}

%===========================================================
\begin{titlepage}
\begin{center}

\textbf{\LARGE Titel der Bachelorarbeit}
%bei langen Titeln, die mehrere Zeilen benoetigen:
%\textbf{\LARGE Langer Titel der  ueber \medskip mehrere Zeilen laeuft}

\bigskip\bigskip
\textbf{Bachelorarbeit}
%bei Masterarbeit:
%\textbf{Masterarbeit}

\bigskip\bigskip\bigskip
Vorgelegt von

\bigskip
\textbf{Vorname Nachname}

\bigskip
aus Geburtsort

\vfill
Angefertigt am\\
Mathematischen Institut\\ 
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\ 
der Heinrich-Heine-Universit\"at D\"usseldorf

\bigskip
%Abgabedatum:
tt.\ Monat jjjj

\bigskip
Betreuer: Prof.\ Dr.\ Vorname Nachname
%Wird der Schwerpunkt der Abschlussarbeit im Anwendungsfach gewaehlt:
%Betreuer: Prof.\ Dr.\ Vorname Nachname\\
%Zweitbetreuer: Prof.\ Dr.\ Vorname Nachname

\end{center}
\end{titlepage}

\thispagestyle{empty}\mbox{}\pagebreak
\setcounter{page}{0}

%===========================================================
\tableofcontents
\pagebreak

%===========================================================
\section*{Einleitung}
\addcontentsline{toc}{section}{Einleitung}

In der vorgelegten Abschlussarbeit untersuchen wir ...

\pagebreak

%===========================================================
\section{Task 3}
Focus tracker accumulates evidence of a changing state over the course of the dialogue. Let's define for methods and goal components $c \in V_c$ (all possible slot values) the quantity denoting that a slot(method) is not present in turn $t$:

\begin{equation}
    \label{eqn:q_c_t}
    q_{c,t} := slu_{c,t}('none') = 1 - \sum_{v \in V_c \backslash \{'none'\}}slu_{c,t}(v)
\end{equation}

\noindent where $slu(v)$ stands for 'spoken dialog undertanding' and denotes a probability of a given slot value $v$ of a goal $c$ being the correct one.\\

\noindent Let's recursively define probabilities $p_{c,t}(v)$ for each $v$ in $V_c \backslash \{'none'\}$ as:

\begin{align}
    \label{eqn:p_c_t}
    p_{c,t}(v) &=slu_{c,t}(v) + q_{c,t} \cdot p_{c,t-1}(v) \\
    \label{eqn:p_c_0}
    p_{c,0}(v) &= slu_{c,0}(v) \\
    \label{eqn:p_c_t_none}
    p_{c,t}('none') &= 1 - \sum_{v \in V_c \backslash \{'none'\}}p_{c,t}(v)
\end{align}

\noindent For a requestable slot $s$ we define 
\begin{align}
    \label{eqn:p_s_t}
    p_{s,t} &= slu_{s,t} + (1 - slu_{s, t})\cdot p_{s,t-1} \\
    \label{eqn:p_s_0}
    p_{s,0} &= slu_{s,0}
\end{align}

\noindent Let's prove that probabilities are well-defined, in other words that $0 \leq p_{c,t}(v) \leq 1$ $\forall v \in V_c, \forall t, \forall c$

\begin{proof}
    For goals(methods).
    Using method of mathematical induction for $t=0$: \\
    From \ref{eqn:p_c_0} since $0 \leq slu_{c,t}(v) \leq 1 \implies $
    \begin{equation}
        \begin{aligned}
            0 \leq &p_{c,0}(v) \leq 1 \\
            0 \leq &p_{c, 0}('none') \leq 1
        \end{aligned}
    \end{equation}

    \noindent Assuming that probabilities are well-defined for $t = k-1$:
    \begin{equation}
        0 \leq p_{c,k-1}(v) \leq 1
    \end{equation} 
    Since
    \begin{equation}
        \begin{aligned}
            0 \leq slu_{c,k}(v) + slu_{c,k}('none') = 1 - \sum_{w \in V_{c} \backslash \{'none', v\}} slu_{c,k}(w) \leq 1
        \end{aligned}
    \end{equation}
    and
    $ slu_{c,k}(v) \geq 0 $,  $q_{c,k} \geq 0$, $p_{c,k-1}(v) \geq 0$
    then for $t=k$ we get
    \begin{equation}
        \begin{aligned}
            0 \leq p_{c, k}(v) \overset{
                \mathrm{\ref{eqn:p_c_t}}, 
                \mathrm{\ref{eqn:q_c_t}}
                }{=} slu_{c,k}(v) + q_{c,k} \cdot p_{c, k-1}(v) \leq 1
        \end{aligned}
    \end{equation}
    As $p_{c, t}('none') \leq 1$ by definition (\ref{eqn:p_c_t_none}), we only need to prove it being $\geq 0$.\\

    \noindent Again by method of mathematical induction let's assume that for $t = k-1$ $p_{c, k-1}('none') \geq 0$ then for $t = k$

    \begin{equation}
        \begin{aligned}
            p_{c,k}('none') 
            &\overset{
                \mathrm{\ref{eqn:p_c_t_none}},
                \mathrm{\ref{eqn:p_c_t}}, 
                \mathrm{\ref{eqn:q_c_t}}
                }{=} 1 - \sum_{v \in V_{c} \backslash \{'none'\}} \left[ slu_{c,k}(v) + q_{c,k} \cdot p_{c,k-1}(v) \right] \\ 
            &\overset{
                \mathrm{\ref{eqn:q_c_t}}
                }{=} slu_{c,k}('none') - \sum_{v \in V_{c} \backslash \{'none'\}} slu_{c,k}('none') \cdot p_{c,k-1}(v) \\ 
            &= slu_{c,k}('none')(1 - \sum_{v \in V_{c} \backslash \{'none'\}}p_{c, k-1}(v)) 
            \\
            &\overset{\sum_{V_c}p_{c, k-1}(v) = 1}{=} slu_{c,k}('none') \cdot p_{c, k-1}('none') \geq 0
        \end{aligned}
    \end{equation}

    For requestable slots.
    Assuming that probabilities are well-defined for $t = k-1$:
    \begin{equation}
        0 \leq p_{s,k-1} \leq 1
    \end{equation}
    Since
    \begin{equation}
        0 \leq slu_{s,t} + (1 - slu_{s,t}) \leq 1
    \end{equation}

    \begin{equation}
        0 \leq p_{s,t} \overset{\ref{eqn:p_s_t}}{=} slu_{s,t} + (1 - slu_{s,t}) \cdot p_{s,t-1} \leq 1
    \end{equation}
\end{proof}

\subsection{Exercise 4}

$slu = slu_{c, 1}(v) = slu_{c,2}(v)$ for $\forall c \forall v \in V_c \Rightarrow q_{c,1} = q_{c, 2} = q$

\begin{equation}
    \begin{aligned}
        &p_{c,2}(v) = slu_{c,2}(v) + q_c \cdot p_{c, 1}(v) 
        \\
        &p_{c,1}(v) = slu_{c,1}(v)
        \\
        &p_{c, 2}(v) = slu_{c,2}(v) + q_c \cdot p_{c, 1}(v) = slu(v)(1 + q_c)
    \end{aligned}
\end{equation}

$\Rightarrow$ probability of value $v$  for slot $c$ will grow $p_{c,2}(v) \geq p_{c,1}(v) \forall v \in V_c $

\begin{equation}
    \begin{aligned}
        p_{c,2}('none') = 1 - \sum_{v \in V_{c} \backslash \{'none'\}} p_{c,2}(v) \leq 1 - \sum_{v \in V_{c} \backslash \{'none'\}} p_{c,1}(v)  = p_{c,1}('none')
    \end{aligned}
\end{equation}

So the probability of value 'none' being 'none' when it is repeated twice decreases.

\subsection{Exercise 5}

\begin{equation}
    p_{c,t}(v) = (1-q_{c,t})\cdot slu_{c,t}(v) + q_{c,t} \cdot p_{c,t-1}(v)
\end{equation}

Since $q_{c,0}=1$ means that all ptobabilities were 0 for any $v$ apart from 'none' value, so $p_{c,0}(v) = 0$ for all $v \in V_c$

Then $p_{c,1}(v) = (1 = 0.5) \cdot 0.5 + 0 = 0.25$

As the slot was not in the first turn and then appeared in the second turn with probability 0.5, then it is okay to assume that the the probability of it might be not 0.5 (as it would be in previous update rule) but 0.25.

In general, such rule puts less weights on the new probabilities from current turn and puts more weights on the previouos turn than before.

\subsection{Exercise 3}
it performes much better for the informable slots. This is an expected behaviour as informable slot may change over time, and the probability of the slot being informable will change over time. However for the requestable slot focus tracker doesn't show a better performance and usually the requestable slots do not change during the dialog, they are requested once and after that removed from the set of the requestable slots. The same applied for the methods, performance there is the similar to the baseline tracker as user only once "sets" by which method he wants to achieve the goal.
\section{Monte-Carlo policy}

Monte Carlo control algorithm is a method for finding the optimal policy for a reinforcement problem.

It is a tabular approach, which means that the belief state/summary space is discretised into a grid. In this algorithm the policy and the Q function are first defined randomly. Then we iterate over the dialogs with this polilcy. Throught this iteration we calculate the reward until the end of the dialog. For updating the state we choose the nearest state from the table and update the Q value and afterwards the policy with it, then in the next turn we will use an already updated policy.

\begin{center}
    \includegraphics[width=10cm]{MCC.png}
\end{center}
It is a value-based algorithm, which means that it is based on the value of the state-action pairs. It is also a model-free method as it does not require any knowledge of the environment. And finally it is a on-policy as we use the current best estimate of the policy to guide the agent. And policy is updated iteratively until we find the optimal one. Regarding the exploration, it is a greedy method as we sometimes choose the action with the highest value, and sometimes pick a random one.

\subsection{Why do we define a reward?}

Reward is a measure of the success of the policy for the dialog, if we don’t have a reward, we cannot learn and define an optimal policy. With it we can define how we optimize the policy by optimizing the value of the reward. It is basically a metric for the algorithm's success.

\subsection{What actions can system execute?}

Interesting detail for optimizing the training process is explicitly define which actions can the system execute. This is important because we want to train the system to execute only those actions that are possible. For example, if the system is in a state where it cannot execute the action “say hello” (after user requests a restaurant's pnone number), then we should not train the system to execute that action.

In order to implement that we can use a mask to define which actions can be executed. The mask is a binary vector of size equal to the number of actions. If the mask is 1, then the action is allowed to be executed. If the mask is 0, then the action is not allowed to be executed.

\pagebreak

%===========================================================
\section{Deep Q-Network}

%===========================================================
\subsection{Q-Learning}
Instead of keeping Q-values in a table, we keep them in a network. This network is called a Q-network. The Q-network is a neural network that takes the state of the dialog as input and outputs the Q-value for each of possible actions. The optimal Q-function (Q-network) obeys the following Bellman optimality equation:

\[ Q*(s,a) = \E [r + \gamma max_{a^\prime}Q(s^\prime,a^\prime)] \]

where $\gamma$ is the discount factor,  $r$ is the immediate reward and $\max_a Q(s',a)$ is the maximum Q-value of the next state. The idea of the DQN paper is to use this Bellman optimality equation to update the Q-function in an interative way:
\[Q_{i + 1} (s,a) \leftarrow \E[r + \gamma max_{a^\prime}Q_{i}(s^\prime,a^\prime)]\]
It can be shown that with these iterations we will converge to the optimal Q-function.

In comparison to Monte-Carlo approach this one is more efficient because it does not need to keep track of all the past states. Instead it only needs to keep track of the current state. This is because the Q-value of the current state is used to calculate the Q-value of the next state.

DQN approach benefints from two main ideas: experience replay and delayed Q-targets. For this type of learning the tuples of ${s, a, r, s^\prime}$, there $s^\prime$ is the following dialog state. The loss which will be minimized is the difference between the Q-value of the current state and the Q-value of the next state:

\[\L = \E_{(s, a, r, s^\prime) \sim D} [(Q_{target}(r,s^\prime) - Q(s, a))^2]\]

where $Target(r, s^\prime) = r + \gamma max_{a}Q_{target}(s^\prime,a)$

and if $s^\prime$ is a terminal state then $Q(s^\prime, a) = Q_{target}(s^\prime, a) = 0$ for every $a$.

As well as with Monte-Carlo we choose and $e$-greedy action selection. In this case training with the second environment is much more efficient and accurate, than training the same environment using Monte-Carlo approach.

Experience replay provides two main advantages: first as we store data tuples $(s, a, r, s^\prime)$ in an experience buffer $D$, then each step of the experience that model encounters is potentially used in many weight updates. Second, as samples are very correlated, learning from subsequent samples is not efficient and creates a high variance.

Delayed Q-Targets means that the target network is updated with the delay, which stabilizes the network even more. The updates of the target network work as a mean-time-average?????

Additionally we have trained a Q-network without the experience replay. In this case the result was much worse and the variance was much higher, therefore the network was very unstable.

\begin{center}
    \begin{tabular}{||c c c c||} 
     \hline
      & Reward & Success & Turns \\ [0.5ex] 
     \hline\hline
     Experience replay & 11.23 +- 0.97 & 93.00 +- 3.56 & 7.37 +- 0.53 \\ 
     \hline
     No experience replay & 4.50 +- 2.04 & 69.00 +- 6.45 & 9.30 +- 0.94 \\
     \hline
    \end{tabular}
\end{center}

The same informations is presente also in the plots.

\begin{figure}[!htb]
    \minipage{0.32\textwidth}
      \includegraphics[width=\linewidth]{env2-CamRestaurants-experience.png}
      \caption{Experience Replay}
    \endminipage\hfill
    \minipage{0.32\textwidth}
      \includegraphics[width=\linewidth]{env2-CamRestaurants-experience.png}
      \caption{No Experience Replay}
    \endminipage\hfill
    \minipage{0.32\textwidth}%
      \includegraphics[width=\linewidth]{env2-CamRestaurants-mc.png}
      \caption{Monte-Carlo}
    \endminipage
\end{figure}

As the system is not supposed to live on its own, here we present several dialogs conducted with it. 
%=========================================================== 
\section{Zweiter Abschnitt}

However, these methods were previously quite unstable (Mnih et al., 2013). In DQN,
Mnih et al. (2013, 2015) proposed two techniques
to overcome this instability-namely experience replay and the use of a target network. In experience replay, all the transitions are put in a finite
pool D (Lin, 1993). Once the pool has reached
its predefined maximum size, adding a new transition results in deleting the oldest transition in
the pool. During training, a mini-batch of transitions is uniformly sampled from the pool, i.e.
(Bt
, At
, Rt+1, Bt+1) from U(D). This method removes the instability arising from strong correlation between the subsequent transitions of an
episode (a dialogue). Additionally, a target network with weight vector w is used. This target
network is similar to the Q-network except that
its weights are only copied every tau steps from the
Q-network, and remain fixed during all the other
steps. The loss function for the Q-network at iter

\pagebreak
\begin{thebibliography}{ccccc}
\addcontentsline{toc}{section}{Literatur}

%So kann eine eine Monographie zitieren:
\bibitem{Arbarello et al 1985}
E.\ Arbarello, M.\ Cornalba, P.\ Griffiths, J.\ Harris:
Geometry of algebraic curves. I. 
Springer, New York, 1985.

%So kann man eine Originalarbeit zitieren:
\bibitem{Shepherd-Barron 1997}
N.\ Shepherd-Barron:
Fano threefolds in positive characteristic.
Compositio Math.\  105  (1997),  237--265.

\end{thebibliography}

%===========================================================
\pagebreak\noindent
\textbf{\LARGE Erkl\"arung}
\addcontentsline{toc}{section}{Erkl\"arung}

\bigskip\bigskip
\noindent 
Hiermit versichere ich, dass ich die   Bachelorarbeit selbst\"andig verfasst und keine
anderen als die angegebenen Quellen und Hilfsmittel benutzt habe.

\bigskip
\noindent
D\"usseldorf, den tt.\ Monat jjjj

\bigskip\bigskip\bigskip
\noindent
(Vorname Nachname)

\end{document}

### 2021-12-26 paper.tex
task 2

@@
 \pagebreak
+\section{Task 2 Semantic Decoding}
+Semantic decoding translates sentences into the set of sematic concepts, which consists of 2 parts: intent (or dialog act type) and sematic slots and values.
+For instance the sentence "I want to book a flight to Berlin" can be translated into the intent "book" and the slots "flight" and "to" with the value "Berlin"($book(flight, to(Berlin))$).
 
+To do this we first need to translate a sentence into a sequence of tags, for ex.
+"I want to fly from Minsk to Kiev round trip tomorrow" can be translated into the sequence of tags: ["O", "O", "O", "O", "O", "B-fromloc.city\_name", "O", "B-toloc.city\_name", "B-round\_trip", "I-round\_trip", "B-departure\_date"].
+
+In order to do this, we can keep two dictionaries (one for words and one for tags), which will map each word (tag) to a unique integer.
+We can then use these dictionaries to translate the sequence of words and a sequence of tags into a sequence of integers.
+Then we can train a model to predict from the sequence of integers a corresponding sequence of tags.
+This can be a simple RNN with two heads - one for predicting the tags and another for predicting intent.
+
+There are different ways of how we can evaluate such model.
+But first, having a sequence of tags, we can convert it to a set of slots in the following way: input = [B-food-type I\_food-type O B\_price-range O], output = [(food-type, start=0, end=1), (price-range, start=3, end=3)].
+Then having a set of true and predicted slots we can compare them on the slot level: if predicted slot was in true slots, then this is a true positive.
+If predicted slot was not in true slots, then this is a false positive.
+If true slot was not in predicted slots, then this is a false negative.
+Then from these values, we can calculate precision, recall and F1 score.
+
+On the other hand, we can calculate accuracy directly on the predicted sequence of tags, without converting it to slots.
+In this case we keep the mapping of each tag to 3 counters for TP, FP and FN.
+After that there are 2 options on how to average the metrics.
+On a macro level, we first calculate precision, recall and f1-score for each of the tags, then we average these metrics across all tags.
+On a micro level we do the weighted average, so for ex.
+to calculate precision we divide the sum of all TP by the sum of all TP+FP of all tags.
+
+Macro average can be much lower than a weighted average, because there can be a lot of rare classes with few instances which were not predicted or predicted wrongly while there are not so many classes with a lot of instances, which are therefore predicted correctly.
 %===========================================================
@@
 
-    For requestable slots.
+    For requestable slots.\\
+
     Assuming that probabilities are well-defined for $t = k-1$:
@@
     \end{equation}
+
     Since
+
     \begin{equation}
@@
 
+\subsection{Exercise 3}
+Focus tracker performes much better for the informable slots.
+This is an expected behaviour as informable slot may change over time, and the probability of the slot being informable will change over time.
+However for the requestable slot focus tracker doesn't show a better performance and usually the requestable slots do not change during the dialog, they are requested once and after that removed from the set of the requestable slots.
+The same applied for the methods, performance there is the similar to the baseline tracker as user only once "sets" by which method he wants to achieve the goal.
+
 \subsection{Exercise 4}
+What happens to a probability if a user gives the exact same utterance in
+the first two turns of the dialog?
 
-$slu = slu_{c, 1}(v) = slu_{c,2}(v)$ for $\forall c \forall v \in V_c \Rightarrow q_{c,1} = q_{c, 2} = q$
+$slu = slu_{c, 1}(v) = slu_{c,2}(v) \Rightarrow q_{c,1} = q_{c, 2} = q$ for $\forall c \forall v \in V_c $
 
@@
 
-So the probability of value 'none' being 'none' when it is repeated twice decreases.
+So the probability of value 'none' when the same utterance is repeated twice decreases.
 
 \subsection{Exercise 5}
-
+Let's defince another update rule:
 \begin{equation}
@@
 \end{equation}
+Assume we talked to the user for two turns and for an informable slot $c$ it holds that 
+\begin{equation}
+    \label{eqn:q_c_0_1}
+    q_{c, 0} = 1
+\end{equation}
+and 
+\begin{equation}
+    \label{eqn:q_c_1}
+    q_{c, 1} = slu_{c,1}(v) = 0.5
+\end{equation}
 
-Since $q_{c,0}=1$ means that all ptobabilities were 0 for any $v$ apart from 'none' value, so $p_{c,0}(v) = 0$ for all $v \in V_c$
+\noindent What happens to the probability $p_{c,1}(v)$?\\
 
-Then $p_{c,1}(v) = (1 = 0.5) \cdot 0.5 + 0 = 0.25$
+\noindent Since (\ref{eqn:q_c_0_1}) means that all probabilities were 0 for any $v$ apart from 'none' value, so $p_{c,0}(v) = 0$ for all $v \in V_c$ \\
 
-As the slot was not in the first turn and then appeared in the second turn with probability 0.5, then it is okay to assume that the the probability of it might be not 0.5 (as it would be in previous update rule) but 0.25.
+\noindent Then $p_{c,1}(v) = (1 = 0.5) \cdot 0.5 + 0 = 0.25$ \\
 
-In general, such rule puts less weights on the new probabilities from current turn and puts more weights on the previouos turn than before.
+\noindent As the slot was not in the first turn and then appeared in the second turn with probability 0.5, then it is okay to assume that the the probability of it might be not 0.5 (as it would be in previous update rule) but 0.25.
+\\
 
-\subsection{Exercise 3}
-it performes much better for the informable slots.
-This is an expected behaviour as informable slot may change over time, and the probability of the slot being informable will change over time.
-However for the requestable slot focus tracker doesn't show a better performance and usually the requestable slots do not change during the dialog, they are requested once and after that removed from the set of the requestable slots.
-The same applied for the methods, performance there is the similar to the baseline tracker as user only once "sets" by which method he wants to achieve the goal.
+\noindent In general, such rule puts less weight on the new probabilities from current turn and more weight on the previous turn, which is also a totally reasonable update.
+
 \section{Monte-Carlo policy}
@@
 
-It is a tabular approach, which means that the belief state/summary space is discretised into a grid.
+\noindent It is a tabular approach, which means that the belief state/summary space is discretised into a grid.
 In this algorithm the policy and the Q function are first defined randomly.

### 2021-12-27 paper.tex
intorduction start

@@
 \usepackage{amsthm}
+\usepackage{hyphenat}
 
@@
 \pagebreak
+\section{Spoken Dialog System}
+A spoken dialogue system is a computer system that enables human computer interaction where primary input is speech.
+Examples of such systems are voice assistants, chatbots, and conversational interfaces (Siri, Alexa, Google Home, etc.).
+\par
+
+A good spoken dialogue system has the following properties: understands the user, decides what to say back, conducts a conversation beyond simple vo\-ice commands or question answering. 
+In this overview we will be considering a goal-oriented systems (the ones which are targeted to accomplish a goal in the context of a task - to book a restaurant for ex.).
+In this case the scope of the conversation is usually limited to a specific domain - a database contating all knowledge about the topic.
+Such database is called an ontology.
+\pagebreak
 \section{Task 2 Semantic Decoding}
@@
 Macro average can be much lower than a weighted average, because there can be a lot of rare classes with few instances which were not predicted or predicted wrongly while there are not so many classes with a lot of instances, which are therefore predicted correctly.
+
+Let's have a look on how the tags are recognised after training. 
+\begin{center}
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{ c c c c c c c c c c c c c }
+        & find & a & flight & from & tampa & to & montreal & by & way & of & new & york
+        \\ 
+        predicted & O & O & O & O & B-fromloc.city\_name & O & B-toloc.city\_name & O & O & O & B-fromloc.city\_name & I-fromloc.city\_name
+        \\ 
+        true & O & O & O & O & B-fromloc.city\_name & O & B-toloc.city\_name & O & O & O & B-stoploc.city\_name & I-stoploc.city\_name \\  
+    \end{tabular}}
+\end{center}
+
+We can fix it by adding words "with a stop in".
+\begin{center}
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{ c c c c c c c c c c c c c c}
+        & find & a & flight & from & tampa & to & montreal & with & a & stop & in & new & york
+        \\ 
+        predicted & O & O & O & O & B-fromloc.city\_name & O & B-toloc.city\_name & O & O & O & O & B-stoploc.city\_name & I-stoploc.city\_name  
+    \end{tabular}}
+\end{center}
+
+There are also many instances when the state name is recognized as a city name. 
+\begin{center}
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{ c c c c c c c c c c c}
+        & please & find & a & flight & from & las & vegas & to & michigan
+        \\ 
+        predicted & O & O & O & O & O & B-fromloc.city\_name & I-fromloc.city\_name &  O & B-toloc.city\_name
+        \\ 
+        true & O & O & O & O & O & B-fromloc.city\_name & I-fromloc.city\_name &  O & B-toloc.state\_name \\  
+    \end{tabular}}
+\end{center}
+
+Arrival time can be mistaken with departing time:
+\begin{center}
+    \scalebox{0.7}{
+    \begin{tabular}{ c c c c c}
+        Sentence & True & Predicted & Changes & New prediction\\
+        i &O&O & i & O\\
+        would &O&O & would & O\\
+        like  &O&O & like & O\\
+        to &O& O & to & O\\
+        return &O& O & fly& O\\
+        from &O& O & around & B-depart\_time.time\_relative\\
+        chicago &B-fromloc.city\_name& B-fromloc.city\_name & 7 & B-depart\_time.time\\
+        around &B-depart\_time.time\_relative& B-depart\_time.time\_relative & in & O\\
+        7 &B-depart\_time.time& B-depart\_time.time & the & O\\
+        pm &I-depart\_time.time& I-arrive\_time.time & evening & B-depart\_time.period\_of\_day\\
+        to &O& O & from & O\\
+        kansas  &B-toloc.city\_name& B-toloc.city\_name & chicago & B-fromloc.city\_name\\
+        city &I-toloc.city\_name& I-toloc.city\_name & to & O \\
+        &&&kansas & B-toloc.city\_name\\
+        &&& city & I-toloc.city\_name\\
+    \end{tabular}}
+\end{center}
+
+We can fix wrongly recognized "pm" by adding - in the evening.
 %===========================================================

### 2021-12-27 paper.tex
diagram

@@
 Such database is called an ontology.
+\par
+One of the straightforward and old approaches to create such a system is rule-base model.
+However such models of course require to handcraft a lot of rules, don't keep track of what has already happend in the dialog and therefore are not prone to annoying repetitive behaviour.
+And here machine learning approaches come in handy.
+\par
+
+System is usually split into modules each responsible for a specific task with well defined input and output, for each a machine learning algorithm can be used.
+Spoken language understanding (SLU), Dialogue management (DM), Spoken language generation (SLG).
+\par
+\includegraphics[width=\linewidth]{SDS-main.png}
+
 \pagebreak

### 2021-12-29 paper.tex
sematic decoding

@@
 Spoken language understanding (SLU), Dialogue management (DM), Spoken language generation (SLG).
+Modular approach has a downside of loosing information between the modules, for example if we have recognised speech incorrectly all following steps will be failed, however ML approaches allow to consider a list of possible inputs scored by their probability, and therefor uncertainty is propagated through the pipeline.
 \par
@@
 
+Both semantic decoding and natural language generation depend on the ontology. Dialogue manager in addition needs to query the ontology to be able to provide an adequate response.
+
+\subsection{Word represeantions}
+To match user utterances with the ontology concepts we either have to either handcode all of the possible user utterances (which is completely uproductive in the constantly changing language and doesn't scale well), or learn their representation implicitely.
+For instance using a distri\-butional approach by encoding words into vectors, which can be learned statistically from the corpus of many texts.
+One can learn static representation with such statistical methods as word2vec, GloVe, fastText (uses n-grams) or contextualized neural language models such as ELMo, BERT etc.
+
+\par
+With static word respresentation one can have a meaningfull distance metric, cosine similarity for ex.: the closer the word vectors are - the more similar meaning they have.
+
+\begin{equation}
+    cos(w_i,w_j) = \frac{w_i^T \cdot w_j}{||w_i|| \cdot ||w_j||}
+\end{equation}
+
+With this approach one can get examples like: $w$("queen") = $w$("king") - $w$("man") + $w$("women")
+\par
+
+\pagebreak
+
+\section{SLU. Speech recognition}
+Speech recognition might be complicated by differences in speakers, environ\-mental noises, divestiry of vocabulary and so on.
+Uncertanties in recognition brought by these factors shoulbe be propagated through the pipeline so that information is not lost after the SLU module because of the wrong recognition.
+\begin{center}
+    \includegraphics{training-1.png}
+\end{center}
+\par 
+There is not straightforward way to convert speech into text - think for example about how to detect when user stars or ends speaking?
+\par Recognition usually doesn't assume the speech being represented in words only, but it works in the hiranhical way: utterance $\Rightarrow$ words $\Rightarrow$ subwords(pho\-nemes, triphones) $\Rightarrow$ 3 senones.
+By subwords we understand the phonemes, by triphones that one can find the in dictionary.
+3 senones represent a subwords unit - specifically its beginning, middle and end.
+
+\begin{center}
+    \includegraphics[width=\linewidth]{hierarcy-speech.png}
+\end{center}
+
+Because we have to align labels in trainging and during recognition the task is to find the most matching sequence from all possible ones we can well apply a HMM model here.
+Each phone is modeled with 3 states.
+The sequence of senones is a sequence of states for HMM.
+Each state of an HMM generates a feature vector of a frame.
+Before pure GMM were applied to solve these problems, afterwards deep learnining came into play hybrid HMM-DNN models appeared.
+There also exist end-to-end transformer models which train from sequence of phones and produce a sequence of frames.
+
 \pagebreak
 \section{Task 2 Semantic Decoding}
-Semantic decoding translates sentences into the set of sematic concepts, which consists of 2 parts: intent (or dialog act type) and sematic slots and values.
+Semantic decoding translates utterances into the set of sematic concepts (dia\-log acts), which consists of 2 parts: intent (or dialog act type) and sematic slots and values.
 For instance the sentence "I want to book a flight to Berlin" can be translated into the intent "book" and the slots "flight" and "to" with the value "Berlin"($book(flight, to(Berlin))$).
 
-To do this we first need to translate a sentence into a sequence of tags, for ex.
+We can view this as a classification task: having probability for each act type and each slot-value pair.
+\begin{center}
+    \includegraphics[width=\linewidth]{training-2.png}
+\end{center}
+Utterances for classification task can be represented as a fixed-size vector with the bag-of-words model.
+Where we preserve only information about which and how many words are there.
+To keep the word order information additionally one can use an n-gram model.
+We can also use delexicalisation to cluster special terms as <area> or <food-type> with the help of ontology in advance.
+Then several SVMs are trained to classify possible dialog acts (see image below).
+\begin{center}
+    \includegraphics[width=\linewidth]{semantic-decoding-classification.png}
+\end{center}
+
+We also consider several possible utterances (see image).
+\begin{center}
+    \includegraphics{uncertanty.png}
+\end{center}
+
+We can view this as a seq-to-seq task: translating an utterance into a sequence of tags, for ex.
 "I want to fly from Minsk to Kiev round trip tomorrow" can be translated into the sequence of tags: ["O", "O", "O", "O", "O", "B-fromloc.city\_name", "O", "B-toloc.city\_name", "B-round\_trip", "I-round\_trip", "B-departure\_date"].
+\begin{center}
+    \includegraphics[width=\linewidth]{training-2-2.png}
+\end{center}
 
@@
 This can be a simple RNN with two heads - one for predicting the tags and another for predicting intent.
+One can also use conditional random fields (CRF) instead.
 

### 2022-01-01 paper.tex
policy as rl

@@
 %===========================================================
-\section{Task 3}
-Focus tracker accumulates evidence of a changing state over the course of the dialogue.
+\section{Dialog state tracking}
+Dialog state tracker remembers everything that has happend in the dialog and that is important for this dialog.
+Dialog management problem can be solved with a generative model with the use of Markov decision process.
+
+\begin{center}
+    \includegraphics[width=\linewidth]{training-state-tr-1.png}
+\end{center}
+
+Process is represented with a dynamic baysian network of states that contain all important information about the dialog so far, actions are what system does and rewards - measures of how good the dialog is so far.
+
+\begin{figure}[!htb]
+    \minipage{0.27\textwidth}
+      \includegraphics[width=\linewidth]{markov.png}
+      \caption{Markov}
+    \endminipage\hfill
+    \minipage{0.26\textwidth}
+      \includegraphics[width=\linewidth]{partial-markov.png}
+      \caption{Partial Markov}
+    \endminipage\hfill
+    \minipage{0.46\textwidth}
+      \includegraphics[width=\linewidth]{state-split.png}
+      \caption{State Split}
+    \endminipage\hfill
+\end{figure}
+
+To introduce the uncertainty here, we can say that the utterances that we observe are not actual states, but that the observations (utterances) we see, and that they are dependent on the state, then the states will be only partially observed and the process is called then partially observable markov process and there is a tractable approximations for it called HIS.
+It works with n-best list of utterances(observed states) and decomposes a state into user goal, user action and dialog history parts.
+
+In generative models we assume that the features between subsequent turns are conditionally independent given the underlying states, but we can model directly dependence of states on the observed features which can be correlated between the turns.
+This approach is called discriminative model.
+
+The dialogue state then consists of three components - goal (for each slot a distribution over possible values), method (a distribution over possible methods), and requested slots (for each requestable slot a binary distribution - wether it was requested or not).
+\par
+Simple tracker might take the most probably state, but one can build a cocus tracker that accumulates evidence of a changing state over the course of the dialogue.
 Let's define for methods and goal components $c \in V_c$ (all possible slot values) the quantity denoting that a slot(method) is not present in turn $t$:
@@
 
-\subsection{Exercise 3}
+\subsection{Exercise 3 Quality Evaluation}
+To evaluate the quality of a belief state tracker one can use accuracy as a fraction of turns where the top dialogue state hypothesis is correct or an l2-norm of the hypothesised distribution $p$ and the true label, where $i$ is index of the true label.
+\begin{center}
+    \begin{equation}
+        L_2 = (1-p_i)^2 + \sum_{j \neq i} p_j^2
+    \end{equation}
+\end{center}
 Focus tracker performes much better for the informable slots.
-This is an expected behaviour as informable slot may change over time, and the probability of the slot being informable will change over time.
+This is an expected behaviour as informable slot may change over time, and the probabi\-lity of the slot being informable will change over time.
 However for the requestable slot focus tracker doesn't show a better performance and usually the requestable slots do not change during the dialog, they are requested once and after that removed from the set of the requestable slots.
@@
 
-\section{Monte-Carlo policy}
+\section{Better models}
+It is also possible to skip semantic decoding module and predict dialog acts directly from the transcribed speech.
+One doesn't have to label the utterances in this case, however dexelicalisation is still needed and it is the main accuracy gainer for such methods.
+And it requires a semantic dictionary for all values from ontology, which is also difficult to obtain in practice.
+
+Delexicalision can be done with word-vector embedding, as for.
+example similar words are used in similar contexts as well as antonyms (south, north, etc.).
+However contextual embedding performs even better for state tracking (when the vector embedding of the word depens on the sentence it is used in).
+This can be obtained with the transformer model.
+Self-attention layers in transformers relate different parts of the input sequence to produce a better representation for it.
+Multihead attention can replace delexicalisation as it compares the input sequence to possible slot values.
+With fine-tuned model like BERT (Bidirectional Encoder Representations from Transformers)
+
+\section{Dialog state tracking as a reinforcement learning problem}
+Depending on the output of the state tracker, the agent can learn to perform the correct action.
+This is a reinforcement learning problem, where the agent is rewarded for performing the correct action.
+The agent can also be penalised for performing the incorrect action - like too long dialog for instance.
+It has the policy (its behaviour that) that it has to optimize, rewards from the environment, value function (prdiction of the future reward given current state) and model of the environment.
+
+\subsection{Why do we define a reward?}
+
+Reward is a measure of the success of the policy for the dialog, if we don’t have a reward, we cannot learn and define an optimal policy.
+With it we can define how we optimize the policy by optimizing the value of the reward.
+It is basically a metric for the algorithm's success.
+It can come form real users of from simulated ones.
+
+\subsection{What actions can system execute?}
+
+Interesting detail for optimizing the training process is explicitly define which actions can the system execute.
+This is important because we want to train the system to execute only those actions that are possible.
+For example, if the system is in a state where it cannot execute the action “say hello” (after user requests a restaurant's pnone number), then we should not train the system to execute that action.
+
+In order to implement that we can use a mask to define which actions can be executed.
+The mask is a binary vector of size equal to the number of actions.
+If the mask is 1, then the action is allowed to be executed.
+If the mask is 0, then the action is not allowed to be executed.
+
+\par
+Next to the value funtion there is a Q-funtion, which determines how good it is to take action $a$ in belief state $b$ and then follow a policy $\pi$?
+
+\begin{center}
+    \begin{equation}
+        Q_{\pi}(b, a) = \E[R_t| b_t = b, a_t = a]
+    \end{equation}
+\end{center}
+
+The optimal value of Q is:
+\begin{center}
+    \begin{align}
+        &V_{*}(b) = max_{\pi}V_{\pi}(b) \\
+        &Q_{*}(b, a) = max_{\pi}Q_{\pi}(b, a) = \E[r_{t+1} + \gamma * V_{*}(b_{t+1})| b_t = b, a_t = a] \\
+    \end{align}
+\end{center}
+
+Bellman Optimality equation for Q-function:
+\begin{center}
+    \begin{equation}
+        Q_{*}(b, a) = \sum_{b^\prime, r} p(b^\prime, r | b, a)[r + \gamma * max_{a^\prime} Q_{*}(b^\prime, a^\prime)]
+    \end{equation}
+\end{center}
+
+Then optimal policy is 
+\begin{center}
+    \begin{equation}
+        \pi_{*}(b) = argmax_{a} Q_{*}(b, a)
+    \end{equation}
+\end{center}
+
+\subsection{Monte-Carlo policy}
 
@@
 
-\noindent It is a tabular approach, which means that the belief state/summary space is discretised into a grid.
+It is a tabular approach, which means that the belief state/summary space is discretised into a grid.
 In this algorithm the policy and the Q function are first defined randomly.
@@
 Regarding the exploration, it is a greedy method as we sometimes choose the action with the highest value, and sometimes pick a random one.
-
-\subsection{Why do we define a reward?}
-
-Reward is a measure of the success of the policy for the dialog, if we don’t have a reward, we cannot learn and define an optimal policy.
-With it we can define how we optimize the policy by optimizing the value of the reward.
-It is basically a metric for the algorithm's success.
-
-\subsection{What actions can system execute?}
-
-Interesting detail for optimizing the training process is explicitly define which actions can the system execute.
-This is important because we want to train the system to execute only those actions that are possible.
-For example, if the system is in a state where it cannot execute the action “say hello” (after user requests a restaurant's pnone number), then we should not train the system to execute that action.
-
-In order to implement that we can use a mask to define which actions can be executed.
-The mask is a binary vector of size equal to the number of actions.
-If the mask is 1, then the action is allowed to be executed.
-If the mask is 0, then the action is not allowed to be executed.
 

### 2022-01-14 paper.tex
Task 6

@@
 Regarding the exploration, it is a greedy method as we sometimes choose the action with the highest value, and sometimes pick a random one.
+Distance between distributions is measured with Jensen-Shannon divergence, which is defined in the following way:
 
@@
 \subsection{Q-Learning}
-Instead of keeping Q-values in a table, we keep them in a network.
-This network is called a Q-network.
-The Q-network is a neural network that takes the state of the dialog as input and outputs the Q-value for each of possible actions.
-The optimal Q-function (Q-network) obeys the following Bellman optimality equation:
+Instead of keeping Q-values in a table, one can approximate a Q-function with a neural network (Q-network).
+It is a neural network that takes a state of a dialog as input and outputs Q-values for each possible action.
+An optimal Q-network obeys the following Bellman optimality equation:
 
-\[ Q*(s,a) = \E [r + \gamma max_{a^\prime}Q(s^\prime,a^\prime)] \]
+\[ Q^*(s,a) = \E [r + \gamma max_{a^\prime}Q(s^\prime,a^\prime)] \]
 
-where $\gamma$ is the discount factor,  $r$ is the immediate reward and $\max_a Q(s',a)$ is the maximum Q-value of the next state.
-The idea of the DQN paper is to use this Bellman optimality equation to update the Q-function in an interative way:
+where $\gamma$ is a discount factor,  $r$ is an immediate reward and $\max_a Q(s',a)$ is the maximum Q-value of the next state. One uses Bellman equation to update the Q-function iteratively:
+
 \[Q_{i + 1} (s,a) \leftarrow \E[r + \gamma max_{a^\prime}Q_{i}(s^\prime,a^\prime)]\]
-It can be shown that with these iterations we will converge to the optimal Q-function.
 
-In comparison to Monte-Carlo approach this one is more efficient because it does not need to keep track of all the past states.
-Instead it only needs to keep track of the current state.
-This is because the Q-value of the current state is used to calculate the Q-value of the next state.
+With iterative updates it will converge to an optimal Q-function.
 
-DQN approach benefints from two main ideas: experience replay and delayed Q-targets.
-For this type of learning the tuples of ${s, a, r, s^\prime}$, there $s^\prime$ is the following dialog state.
-The loss which will be minimized is the difference between the Q-value of the current state and the Q-value of the next state:
+In comparison to Monte-Carlo this approach this one is more efficient as it does not keep track of all the past states.
+Instead it only remembers the current one.
+This happens because only Q-value of the current state is used to calculate a Q-value of the next state.
+Monte-Carlo reaches the maximum success rate of 76\% close to the 2000 dialogs, whereas Q-learning reaches the maximum success rate of 96\% after 2000 dialogs and has already 94\% after ~400 dialogs.
+The reward has almost doubled as well.
+
+Two reasons for benefits of DQN over Monte-Caro are: experience replay and delayed Q-targets.
+For experience replay one stores tuples ${s, a, r, s^\prime}$ in the buffer, where $s^\prime$ is the following dialog state and samples batches from the it.
+Delayed targets allow to stabilize the network as we the target network’s parameters are not trained, but they are periodically synchronized with the parameters of the main Q-network.
+The loss funciton here is the difference between Q-value of the current state + current reward vs.
+Q-value of the next state:
 
@@
 
-where $Target(r, s^\prime) = r + \gamma max_{a}Q_{target}(s^\prime,a)$
+where $Q_{target}(r, s^\prime) = r + \gamma max_{a}Q_{target}(s^\prime,a)$
 
-and if $s^\prime$ is a terminal state then $Q(s^\prime, a) = Q_{target}(s^\prime, a) = 0$ for every $a$.
-
-As well as with Monte-Carlo we choose and $e$-greedy action selection.
-In this case training with the second environment is much more efficient and accurate, than training the same environment using Monte-Carlo approach.
-
-Experience replay provides two main advantages: first as we store data tuples $(s, a, r, s^\prime)$ in an experience buffer $D$, then each step of the experience that model encounters is potentially used in many weight updates.
-Second, as samples are very correlated, learning from subsequent samples is not efficient and creates a high variance.
-
-Delayed Q-Targets means that the target network is updated with the delay, which stabilizes the network even more. The updates of the target network work as a mean-time-average?????
+if $s^\prime$ is a terminal state, then $Q(s^\prime, a) = Q_{target}(s^\prime, a) = 0$ for every $a$.
 
@@
 
-As the system is not supposed to live on its own, here we present several dialogs conducted with it. 
+As the system is not supposed to live on its own, here we present several dialogs conducted with it:
+
+\begin{center}
+    \includegraphics[width=\linewidth]{address-1.png}
+\end{center}
+
+\begin{center}
+    \includegraphics[width=\linewidth]{phone.png}
+\end{center}
+
+\section{Information Gain}
+
+One can also substitute overall reward $r$ with $r = r + \nu r_i$, here $r_i$ is an additional intrinsic reward, and $\nu$ is a hyperparameter.
+This reward is called information gain and it measures how much additional information the system gets about user needs after one turn.
+Let action be of form $\{request, confirm,$ $select\}\_\xi$.
+Let $p_{\xi}$ and ${p_{\xi}}^\prime$ be belief distributions over possible values for $\xi$ in current and next belief state.
+Then the informa\-tion gain is the difference between these distributions with Jensen-Shannon divergence as a distance metric:
+
+\begin{equation}
+    r_i(s, a, s\prime) =
+    \begin{cases}
+        d(p_{\xi}, p_{\xi}^\prime) \\
+        0, \text{if } a \notin \text{\{request, confirm, select\}}
+    \end{cases}
+\end{equation}
+where
+\begin{equation}
+    \begin{aligned}
+        &d(p, q) = JS(p,q) = \frac{1}{2}(KL[p||m] + KL[q||m]), \\
+        &m = \frac{1}{2}(p + q)
+    \end{aligned}
+\end{equation}
+
+When the distance is big enough (higher than a threshold) we add the information gain to a reward additionally.
+Reward will be added only when the action was of the form ${request, confirm, select}$.
+Intuitively it also doesn't make sense to calculate information gain for actions like $bye, inform$ as they do not provide any additonal information on user needs and therefore do not lead to the accomplishment of the dialog goal.
+It is very beneficial to use intrinsic rewards in addition as they encourage system to get as much useful information from users as possible, by requesting more information from them, selecting the most relevant one and confirming the choices.
+
+Here one can see the difference in training with information gain and without it.
+Information gain improves the speed of training in the beginning and slightly improves the overall performance from $96\%$  to $98.4\%$ (see Figure 7, 8).
+
+\begin{figure}[!htb]
+    \minipage{0.5\textwidth}
+      \includegraphics[width=\linewidth]{env3-ig-CamRestaurants.png}
+      \caption{Information Gain}
+    \endminipage\hfill
+    \minipage{0.5\textwidth}
+      \includegraphics[width=\linewidth]{env3-CamRestaurants.png}
+      \caption{No Information Gain}
+    \endminipage
+\end{figure}
+
+One can also substitute $r = r + \nu r_i$.
+However one should be cautious --- if the value $\mu$ is too high, we push the sytem to gather more and more information and discourage it from making a choice and informing a user about it.
+
 %=========================================================== 

### 2022-01-14 paper.tex
Build troubles

@@
 This happens because only Q-value of the current state is used to calculate a Q-value of the next state.
-Monte-Carlo reaches the maximum success rate of 76\% close to the 2000 dialogs, whereas Q-learning reaches the maximum success rate of 96\% after 2000 dialogs and has already 94\% after ~400 dialogs.
+Monte-Carlo reaches the maximum success rate of TODO 76\% close to the 2000 dialogs, whereas Q-learning reaches the maximum success rate of 96\% after 2000 dialogs and has already 94\% after ~400 dialogs.
 The reward has almost doubled as well.
@@
 
-Additionally we have trained a Q-network without the experience replay. In this case the result was much worse and the variance was much higher, therefore the network was very unstable.
+Additionally, we have trained a Q-network without experience replay.
+In this case reward was almost 3 times lower TODO (11.23 and 4.5) as well as the success rate.
+The variance has also increased and therefore the network became more unstable.
+One can see the results in Table TODO.
 
@@
      \hline\hline
-     Experience replay & 11.23 +- 0.97 & 93.00 +- 3.56 & 7.37 +- 0.53 \\ 
+     Experience replay & TODO 11.23 +- 0.97 & 93.00 +- 3.56 & 7.37 +- 0.53 \\ 
      \hline
@@
 
+TODO change images
 \begin{figure}[!htb]
@@
 
-As the system is not supposed to live on its own, here we present several dialogs conducted with it:
+Here are several dialogs conducted with the system:
 
 \begin{center}
-    \includegraphics[width=\linewidth]{address-1.png}
+    \includegraphics[width=\linewidth]{wants_food_type.png}
 \end{center}
@@
 \begin{center}
-    \includegraphics[width=\linewidth]{phone.png}
+    \includegraphics[width=\linewidth]{name\ !=.png}
 \end{center}
+
+One can see that system sometimes is caught in the loop and keeps asking the same question over and over again.
+The reason might be that semantic decoder didn't work well and haven't recognised an imput correctly, so the systems continue to request more information.
+TODO get example
+
+It also noticable that system doesn't work well with answers like "I don't care" or "It doesn't matter" and insists on user making a choice (food-type for ex.
+see Figure TODO).
+This might be easily explained by the fact that ontoly doesn't contain such options.
+
+Somtimes it tries to inform on something else instead of providing a phone number.
+It might be the fault of the semantic decoder though, as the system decides that utterance "phone" is not a phone number request, but a request to find another restaurant.
+It might be that longer more human-like sentences "And what is the address of the suggested restaurant?" are more suitable.
 

### 2022-01-16 paper.tex
task 5

@@
 
-\graphicspath{ {./images/} }
+\graphicspath{ {./images} }
 \DeclareMathOperator*{\E}{\mathbb{E}}
@@
 
-It also noticable that system doesn't work well with answers like "I don't care" or "It doesn't matter" and insists on user making a choice (food-type for ex.
-see Figure TODO).
-This might be easily explained by the fact that ontoly doesn't contain such options.
+It also noticable that system doesn't work well with answers like "I don't care" or "It doesn't matter" and insists on user making a choice (food-type for ex. see Figure TODO).
 
@@
 It might be that longer more human-like sentences "And what is the address of the suggested restaurant?" are more suitable.
+
+There was no instances in the dataset where the user would say "I don't care" regarding the price range, and we can see an example of how the dialog is ruined when we encounter such a case, compare dialog 1 vs dialog 2.
+TODO dialog number Usually such infinite loops may happen due to bad recognition of speech, however here the problem is hidden probably in the state tracker.
+
+\begin{figure}[!htb]
+    \minipage{0.49\textwidth}
+      \includegraphics[width=\linewidth]{with_price.png}
+      \caption{With price}
+    \endminipage\hfill
+    \minipage{0.49\textwidth}%
+      \includegraphics[width=\linewidth]{without_price.png}
+      \caption{Without price}
+    \endminipage
+\end{figure}
+
+Another noticable thing is that choice of value for one slot sometimes influences the dialog drastically even though the slot is not that relevant.
+For example, here in the dialog 3 we choose specific price in comparison with dialog 4 TODO check number where we are agreed to any price and that leads suddenly to an infinite loop.
+This might happen due to insufficient semantic decoder knowledge and not enough of such samples in the database.
+\begin{figure}[!htb]
+    \minipage{0.49\textwidth}
+      \includegraphics[width=\linewidth]{chinese_works.png}
+      \caption{Working sample}
+    \endminipage\hfill
+    \minipage{0.49\textwidth}%
+      \includegraphics[width=\linewidth]{chinese_no_work.png}
+      \caption{Not working sample}
+    \endminipage
+\end{figure}
+
+Examples with negations are performing poorly as well which is also most probably the failure of the semantic decoder (see dialog 5 TODO change).
+
+\begin{figure}
+    \includegraphics[width=5cm]{chinese\ negation.png}
+\end{figure}
 

### 2022-01-16 paper.tex
deletions

@@
 
-Here are several dialogs conducted with the system:
-
-\begin{center}
-    \includegraphics[width=\linewidth]{wants_food_type.png}
-\end{center}
-
-\begin{center}
-    \includegraphics[width=\linewidth]{name\ !=.png}
-\end{center}
-
 One can see that system sometimes is caught in the loop and keeps asking the same question over and over again.
@@
 TODO get example
-
-It also noticable that system doesn't work well with answers like "I don't care" or "It doesn't matter" and insists on user making a choice (food-type for ex. see Figure TODO).
-
-Somtimes it tries to inform on something else instead of providing a phone number.
-It might be the fault of the semantic decoder though, as the system decides that utterance "phone" is not a phone number request, but a request to find another restaurant.
-It might be that longer more human-like sentences "And what is the address of the suggested restaurant?" are more suitable.
 

### 2022-01-16 paper.tex
updates task 4

@@
 \section{Dialog state tracking as a reinforcement learning problem}
-Depending on the output of the state tracker, the agent can learn to perform the correct action.
-This is a reinforcement learning problem, where the agent is rewarded for performing the correct action.
-The agent can also be penalised for performing the incorrect action - like too long dialog for instance.
-It has the policy (its behaviour that) that it has to optimize, rewards from the environment, value function (prdiction of the future reward given current state) and model of the environment.
+Apart from tracking a state of a dialog, the agent must learn to perform the correct action.
+This is can be viewed as a reinforcement learning problem, where the agent is rewarded for performing correct actions and might be penalised for the incorrect ones.
+Agent owns a policy (its behaviour) that it has to optimize, rewards from the environment, value function (prediction of the future reward given current state $b$) and a model of the environment.
 
-\subsection{Why do we define a reward?}
+One also defines a Q-function, which determines how good it is to take action $a$ in belief state $b$ and then follow a policy $\pi$.
+Having an optimal Q-function one chooses an action that maximises its value, this will be an optimal action.
 
-Reward is a measure of the success of the policy for the dialog, if we don’t have a reward, we cannot learn and define an optimal policy.
+\begin{equation}
+    Q_{\pi}(b, a) = \E[R_t| b_t = b, a_t = a]
+\end{equation}
+
+The optimal value of Q is:
+\begin{align}
+    &V_{*}(b) = max_{\pi}V_{\pi}(b) \\
+    &Q_{*}(b, a) = max_{\pi}Q_{\pi}(b, a) = \E[r_{t+1} + \gamma * V_{*}(b_{t+1})| b_t = b, a_t = a]
+\end{align}
+
+Optimal policy is: 
+\begin{equation}
+    \pi_{*}(b) = argmax_{a} Q_{*}(b, a)
+\end{equation}
+
+\subsection{Why do to define a reward?}
+
+Reward is a measure of success of the policy for the dialog, if we don’t have a reward, we cannot learn and define an optimal policy.
 With it we can define how we optimize the policy by optimizing the value of the reward.
 It is basically a metric for the algorithm's success.
-It can come form real users of from simulated ones.
 
@@
 
-Interesting detail for optimizing the training process is explicitly define which actions can the system execute.
-This is important because we want to train the system to execute only those actions that are possible.
-For example, if the system is in a state where it cannot execute the action “say hello” (after user requests a restaurant's pnone number), then we should not train the system to execute that action.
-
-In order to implement that we can use a mask to define which actions can be executed.
-The mask is a binary vector of size equal to the number of actions.
-If the mask is 1, then the action is allowed to be executed.
-If the mask is 0, then the action is not allowed to be executed.
-
-\par
-Next to the value funtion there is a Q-funtion, which determines how good it is to take action $a$ in belief state $b$ and then follow a policy $\pi$?
-
-\begin{center}
-    \begin{equation}
-        Q_{\pi}(b, a) = \E[R_t| b_t = b, a_t = a]
-    \end{equation}
-\end{center}
-
-The optimal value of Q is:
-\begin{center}
-    \begin{align}
-        &V_{*}(b) = max_{\pi}V_{\pi}(b) \\
-        &Q_{*}(b, a) = max_{\pi}Q_{\pi}(b, a) = \E[r_{t+1} + \gamma * V_{*}(b_{t+1})| b_t = b, a_t = a] \\
-    \end{align}
-\end{center}
-
-Bellman Optimality equation for Q-function:
-\begin{center}
-    \begin{equation}
-        Q_{*}(b, a) = \sum_{b^\prime, r} p(b^\prime, r | b, a)[r + \gamma * max_{a^\prime} Q_{*}(b^\prime, a^\prime)]
-    \end{equation}
-\end{center}
-
-Then optimal policy is 
-\begin{center}
-    \begin{equation}
-        \pi_{*}(b) = argmax_{a} Q_{*}(b, a)
-    \end{equation}
-\end{center}
+Interesting trick for optimizing a training process is to explicitely define which actions are allowed for execution.
+An agent should be not allowed to execute those actions which are clearly impossible.
+For example, after user requests a restaurant's pnone number an agent should not be allowed to execute an action “say hello”.
+In order to implement that we can mask out with 0 those actions that are not allowed to be executed.
 
@@
 
-The same informations is presente also in the plots.
+The same informations is presented also in the plots.
 
@@
 
-%=========================================================== 
-\section{Zweiter Abschnitt}
-
-However, these methods were previously quite unstable (Mnih et al., 2013). In DQN,
-Mnih et al. (2013, 2015) proposed two techniques
-to overcome this instability-namely experience replay and the use of a target network. In experience replay, all the transitions are put in a finite
-pool D (Lin, 1993). Once the pool has reached
-its predefined maximum size, adding a new transition results in deleting the oldest transition in
-the pool. During training, a mini-batch of transitions is uniformly sampled from the pool, i.e.
-(Bt
-, At
-, Rt+1, Bt+1) from U(D). This method removes the instability arising from strong correlation between the subsequent transitions of an
-episode (a dialogue). Additionally, a target network with weight vector w is used. This target
-network is similar to the Q-network except that
-its weights are only copied every tau steps from the
-Q-network, and remain fixed during all the other
-steps. The loss function for the Q-network at iter
-
 \pagebreak

### 2022-01-17 paper.tex
work on task 3

@@
 \section{Dialog state tracking}
-Dialog state tracker remembers everything that has happend in the dialog and that is important for this dialog.
-Dialog management problem can be solved with a generative model with the use of Markov decision process.
+Dialog state tracker remembers everything important that has happend in the dialog so far. Dialog management problem can be solved with a generative model with the use of Markov decision process.
 
@@
 
-Process is represented with a dynamic baysian network of states that contain all important information about the dialog so far, actions are what system does and rewards - measures of how good the dialog is so far.
+Process is represented with a dynamic baysian network of states that contain all important information about the dialog so far, actions are what system does and rewards - measures of how good is the dialog so far.
 
@@
 
-To introduce the uncertainty here, we can say that the utterances that we observe are not actual states, but that the observations (utterances) we see, and that they are dependent on the state, then the states will be only partially observed and the process is called then partially observable markov process and there is a tractable approximations for it called HIS.
-It works with n-best list of utterances(observed states) and decomposes a state into user goal, user action and dialog history parts.
-
-In generative models we assume that the features between subsequent turns are conditionally independent given the underlying states, but we can model directly dependence of states on the observed features which can be correlated between the turns.
-This approach is called discriminative model.
+To introduce the uncertainty here, we can say that utterances that we observe are not actual states, but observations we see, and that they are dependent on the state, then the states will be only partially observed and the process is called then partially observable markov process.
+Approximate solution for that process works with n-best list of utterances(observed states) and decomposes a state into user goal, user action and dialog history parts.
 
@@
 \par
-Simple tracker might take the most probably state, but one can build a cocus tracker that accumulates evidence of a changing state over the course of the dialogue.
+Most simple tracker might assume that the most probable state as the current state, but one can build a more advanced focus tracker that accumulates evidence of a changing state over the course of the dialogue.
 Let's define for methods and goal components $c \in V_c$ (all possible slot values) the quantity denoting that a slot(method) is not present in turn $t$:
@@
 \subsection{Exercise 5}
-Let's defince another update rule:
+Let's define another update rule:
 \begin{equation}
@@
 \noindent In general, such rule puts less weight on the new probabilities from current turn and more weight on the previous turn, which is also a totally reasonable update.
-
-\section{Better models}
-It is also possible to skip semantic decoding module and predict dialog acts directly from the transcribed speech.
-One doesn't have to label the utterances in this case, however dexelicalisation is still needed and it is the main accuracy gainer for such methods.
-And it requires a semantic dictionary for all values from ontology, which is also difficult to obtain in practice.
-
-Delexicalision can be done with word-vector embedding, as for.
-example similar words are used in similar contexts as well as antonyms (south, north, etc.).
-However contextual embedding performs even better for state tracking (when the vector embedding of the word depens on the sentence it is used in).
-This can be obtained with the transformer model.
-Self-attention layers in transformers relate different parts of the input sequence to produce a better representation for it.
-Multihead attention can replace delexicalisation as it compares the input sequence to possible slot values.
-With fine-tuned model like BERT (Bidirectional Encoder Representations from Transformers)
 
@@
 \end{center}
-It is a value-based algorithm, which means that it is based on the value of the state-action pairs.
-It is also a model-free method as it does not require any knowledge of the environment.
-And finally it is a on-policy as we use the current best estimate of the policy to guide the agent.
-And policy is updated iteratively until we find the optimal one.
-Regarding the exploration, it is a greedy method as we sometimes choose the action with the highest value, and sometimes pick a random one.
-Distance between distributions is measured with Jensen-Shannon divergence, which is defined in the following way:
 

### 2022-01-17 paper.tex
task 3

@@
 \section{Dialog state tracking}
-Dialog state tracker remembers everything important that has happend in the dialog so far. Dialog management problem can be solved with a generative model with the use of Markov decision process.
+Dialog state tracker remembers everything important that has happened in the dialog so far. Dialog management problem can be solved with the use of Markov decision process (Figure \ref{MDP}).
 
-\begin{center}
+\begin{figure}[ht]
     \includegraphics[width=\linewidth]{training-state-tr-1.png}
-\end{center}
+    \caption{MDP}
+    \label{MDP}
+\end{figure}
 
-Process is represented with a dynamic baysian network of states that contain all important information about the dialog so far, actions are what system does and rewards - measures of how good is the dialog so far.
+Process is represented as a dynamic Bayesian network (Figure \ref{Markov}) of 
+\begin{itemize}
+    \item Dialog states --- contain all current important information about the dialog
+    \item Actions --- what system does
+    \item Rewards --- measures of how good is the dialog so far
+\end{itemize}
 
 \begin{figure}[!htb]
-    \minipage{0.27\textwidth}
+    \minipage{0.38\textwidth}
       \includegraphics[width=\linewidth]{markov.png}
       \caption{Markov}
+      \label{Markov}
     \endminipage\hfill
-    \minipage{0.26\textwidth}
+    \minipage{0.38\textwidth}
       \includegraphics[width=\linewidth]{partial-markov.png}
       \caption{Partial Markov}
-    \endminipage\hfill
-    \minipage{0.46\textwidth}
-      \includegraphics[width=\linewidth]{state-split.png}
-      \caption{State Split}
-    \endminipage\hfill
+      \label{PartialMarkov}
+    \endminipage
 \end{figure}
 
-To introduce the uncertainty here, we can say that utterances that we observe are not actual states, but observations we see, and that they are dependent on the state, then the states will be only partially observed and the process is called then partially observable markov process.
-Approximate solution for that process works with n-best list of utterances(observed states) and decomposes a state into user goal, user action and dialog history parts.
+To introduce an uncertainty here, one can assume that observed utteran\-ces are not actual states, but observations, that are dependent on hidden states, then the process is called then partially observable Markov process (Figure \ref{PartialMarkov}).
 
-The dialogue state then consists of three components - goal (for each slot a distribution over possible values), method (a distribution over possible methods), and requested slots (for each requestable slot a binary distribution - wether it was requested or not).
-\par
-Most simple tracker might assume that the most probable state as the current state, but one can build a more advanced focus tracker that accumulates evidence of a changing state over the course of the dialogue.
-Let's define for methods and goal components $c \in V_c$ (all possible slot values) the quantity denoting that a slot(method) is not present in turn $t$:
+Dialogue state consists of a set of distributions for the next components: 
+\begin{itemize}
+    \item Goal --- for each slot a distribution over possible values
+    \item Method --- a distribution over possible methods
+    \item Requested slots --- for each requestable slot a binary distribution
+\end{itemize}
+
+\subsection{Focus Tracker}
+
+The simplest tracker assigns a single hypothesis for each slot at each turn, a more advanced option is a focus tracker.
+It accumulates evidence of a changing state over the course of the dialogue.
+Here we define how it tracks goals, methods, requestable slots.
+
+Define for methods and goal components $c \in V_c$ (all possible slot values) a quantity denoting that a slot or method is not present in turn $t$:
 
@@
 
-\noindent where $slu(v)$ stands for 'spoken dialog undertanding' and denotes a probability of a given slot value $v$ of a goal $c$ being the correct one.\\
+\noindent where $slu(v)$ and denotes a probability of given slot value $v$ for goal $c$ be\-ing a correct one.
 
-\noindent Let's recursively define probabilities $p_{c,t}(v)$ for each $v$ in $V_c \backslash \{'none'\}$ as:
-
+Recursively define probabilities $p_{c,t}(v)$ for each $v$ in $V_c \backslash \{'none'\}$ as: 
 \begin{align}
@@
 
-\noindent For a requestable slot $s$ we define 
+For a requestable slot $s$ define 
 \begin{align}
@@
 
-\noindent Let's prove that probabilities are well-defined, in other words that $0 \leq p_{c,t}(v) \leq 1$ $\forall v \in V_c, \forall t, \forall c$
+Now let's prove that probabilities are well-defined, in other words that $0 \leq p_{c,t}(v) \leq 1$ $\forall v \in V_c, \forall t, \forall c$
 
@@
 
-\subsection{Exercise 3 Quality Evaluation}
-To evaluate the quality of a belief state tracker one can use accuracy as a fraction of turns where the top dialogue state hypothesis is correct or an l2-norm of the hypothesised distribution $p$ and the true label, where $i$ is index of the true label.
-\begin{center}
+\subsection{Quality Evaluation}
+To evaluate a quality of a belief state tracker one can use a fraction of turns where a top dialogue state hypothesis was correct (accuracy) or use the following equa\-tion (\ref{l2}) where $p$ is a hypothesised distribution, $i$ is an index of a true label.
 \begin{equation}
     L_2 = (1-p_i)^2 + \sum_{j \neq i} p_j^2
+    \label{l2}
 \end{equation}
-\end{center}
-Focus tracker performes much better for the informable slots.
-This is an expected behaviour as informable slot may change over time, and the probabi\-lity of the slot being informable will change over time.
-However for the requestable slot focus tracker doesn't show a better performance and usually the requestable slots do not change during the dialog, they are requested once and after that removed from the set of the requestable slots.
-The same applied for the methods, performance there is the similar to the baseline tracker as user only once "sets" by which method he wants to achieve the goal.
 
-\subsection{Exercise 4}
-What happens to a probability if a user gives the exact same utterance in
+In comparison to the simplest tracker focus tracker performs much better for informable slots.
+This is an expected behaviour as informable slot may change over time, and probabi\-lity of the slot being informable will also change.
+However for the requestable slot focus tracker doesn't show a better performance as usually the requestable slots should not change during the dialog, they are requested once and after that removed from the set of the requestable slots, after that the user can decide to request another one.
+The same applies for methods, performance there is similar to the baseline tracker as user only once "sets" by which method he wants to achieve the goal.
+
+One can see the results for both trackers in Table \ref{focus_tracker_results} and Table \ref{baseline_tracker_results}.
+    \begin{table}[ht]
+        \centering
+        \begin{tabular}{|c c c c|}
+        \hline
+                    & Joint Goals & Requested & Method \\
+        \hline
+        Accuracy   & 0.73        & 0.92      & 0.68 \\ 
+        L2         & 0.42        & 0.12      & 0.58 \\
+        \hline
+        \end{tabular}
+        \caption{Focus Tracker}
+        \label{focus_tracker_results}
+    \end{table}
+
+    \begin{table}[ht]
+        \centering
+        \begin{tabular}{|c c c c|}
+        \hline
+                    & Joint Goals   & Requested & Method \\ 
+        \hline
+        Accuracy   & 0.57          & 0.91      & 0.68 \\ 
+        L2         & 0.83          & 0.12      & 0.58  \\
+        \hline
+        \end{tabular}
+        \caption{Simplest Tracker}
+        \label{baseline_tracker_results}
+    \end{table}
+
+\subsection{Interesting Question}
+What happens to a probability when user gives an exact same utterance in
 the first two turns of the dialog?
 
-$slu = slu_{c, 1}(v) = slu_{c,2}(v) \Rightarrow q_{c,1} = q_{c, 2} = q$ for $\forall c \forall v \in V_c $
+In this case $slu(v) = slu_{c, 1}(v) = slu_{c,2}(v) \Rightarrow q_{c,1} = q_{c, 2} = q$ for $\forall c \forall v \in V_c $
 
-\begin{equation}
-    \begin{aligned}
+\begin{align*}
     &p_{c,2}(v) = slu_{c,2}(v) + q_c \cdot p_{c, 1}(v) 
@@
     \\
-        &p_{c, 2}(v) = slu_{c,2}(v) + q_c \cdot p_{c, 1}(v) = slu(v)(1 + q_c)
-    \end{aligned}
-\end{equation}
+    &p_{c, 2}(v) = slu_{c,2}(v) + q_c \cdot p_{c, 1}(v) = slu(v)(1 + q_c) \geq p_{c,1}(v)
+\end{align*}
 
@@
 
-\begin{equation}
-    \begin{aligned}
+\begin{align*}
     p_{c,2}('none') = 1 - \sum_{v \in V_{c} \backslash \{'none'\}} p_{c,2}(v) \leq 1 - \sum_{v \in V_{c} \backslash \{'none'\}} p_{c,1}(v)  = p_{c,1}('none')
-    \end{aligned}
-\end{equation}
+\end{align*}
 
-So the probability of value 'none' when the same utterance is repeated twice decreases.
+\noindent and the probability of value $none$ decreases.
+This makes an intuitie sense, if a user repeats a sentence an agen should increase a probability of the slots mentioned by him and descrease a probability of $none$ values as user definitely wants to pass to the agent some important information.
 
-\subsection{Exercise 5}
+\subsection{Another update rule}
 Let's define another update rule:
@@
 \end{equation}
+for some $v$.
 
-\noindent What happens to the probability $p_{c,1}(v)$?\\
+What happens to probability $p_{c,1}(v)$?
 
-\noindent Since (\ref{eqn:q_c_0_1}) means that all probabilities were 0 for any $v$ apart from 'none' value, so $p_{c,0}(v) = 0$ for all $v \in V_c$ \\
+Since (\ref{eqn:q_c_0_1}) means that all probabilities were 0 for any $v$ apart from $none$, then $p_{c,0}(v) = 0$ for all $v \in V_c$. Then: 
+\begin{align*}
+    p_{c,1}(v) = (1 - 0.5) \cdot 0.5 + 0 = 0.25
+\end{align*}
 
-\noindent Then $p_{c,1}(v) = (1 = 0.5) \cdot 0.5 + 0 = 0.25$ \\
+As the slot was not in the first turn and then appeared in the second turn with probability 0.5, then it is okay to assume that the the probability of it might be not 0.5 (as it would be in previous update rule) but 0.25.
 
-\noindent As the slot was not in the first turn and then appeared in the second turn with probability 0.5, then it is okay to assume that the the probability of it might be not 0.5 (as it would be in previous update rule) but 0.25.
-\\
+In general, such update rule puts less weight on the new probabilities from current turn and more weight on the previous turn, which is also a totally reasonable update.
 
-\noindent In general, such rule puts less weight on the new probabilities from current turn and more weight on the previous turn, which is also a totally reasonable update.
+\newpage
 

### 2022-01-18 paper.tex
task 2

@@
 \pagebreak
-\section{Task 2 Semantic Decoding}
-Semantic decoding translates utterances into the set of sematic concepts (dia\-log acts), which consists of 2 parts: intent (or dialog act type) and sematic slots and values.
-For instance the sentence "I want to book a flight to Berlin" can be translated into the intent "book" and the slots "flight" and "to" with the value "Berlin"($book(flight, to(Berlin))$).
+\section{Semantic Decoding}
+Semantic decoding translates user utterances into a set of sematic concepts (dia\-log acts).
+Dialog act consists of 2 parts: intent (dialog act type) and sematic slots with their values.
+For instance the sentence "Is there a cheap restaurant in the center?" can be translated into $inform(price=cheap,$ $area$$=center)$.
+Intent is $inform$, slot $price$ with value $cheap$ and slot $area$ with the value $center$.
 
-We can view this as a classification task: having probability for each act type and each slot-value pair.
-\begin{center}
+We can view semantic decoding as a classification task: having a probabi\-lity for each act type and each slot-value pair.
+
+\begin{figure}[h!]
     \includegraphics[width=\linewidth]{training-2.png}
-\end{center}
-Utterances for classification task can be represented as a fixed-size vector with the bag-of-words model.
-Where we preserve only information about which and how many words are there.
+\end{figure}
+
+Utterances for classification task can be represented as a fixed-size vector, where we preserve only information about which and how many words are there.
 To keep the word order information additionally one can use an n-gram model.
-We can also use delexicalisation to cluster special terms as <area> or <food-type> with the help of ontology in advance.
-Then several SVMs are trained to classify possible dialog acts (see image below).
-\begin{center}
-    \includegraphics[width=\linewidth]{semantic-decoding-classification.png}
-\end{center}
+Delexicalisation is used in advance to cluster special terms as $area$ or $food-type$ with the help of ontology.
+Then several SVMs are trained to classify possible dialog acts (Figure \ref{fig:semantic-decoding-classification}).
 
-We also consider several possible utterances (see image).
-\begin{center}
+\begin{figure}[h!]
+    \includegraphics[width=0.998\textwidth]{semantic-decoding-classification.png}
+    \label{fig:semantic-decoding-classification}
+\end{figure}
+
+Several possible utterances are considered (Figure \ref{several}).
+\begin{figure}[h!]
+    \centering
     \includegraphics{uncertanty.png}
-\end{center}
+    \caption{Several Utterances}
+    \label{several}
+\end{figure}
 
@@
 "I want to fly from Minsk to Kiev round trip tomorrow" can be translated into the sequence of tags: ["O", "O", "O", "O", "O", "B-fromloc.city\_name", "O", "B-toloc.city\_name", "B-round\_trip", "I-round\_trip", "B-departure\_date"].
-\begin{center}
+\begin{figure}[h!]
     \includegraphics[width=\linewidth]{training-2-2.png}
-\end{center}
+\end{figure}
 
-In order to do this, we can keep two dictionaries (one for words and one for tags), which will map each word (tag) to a unique integer.
-We can then use these dictionaries to translate the sequence of words and a sequence of tags into a sequence of integers.
-Then we can train a model to predict from the sequence of integers a corresponding sequence of tags.
-This can be a simple RNN with two heads - one for predicting the tags and another for predicting intent.
-One can also use conditional random fields (CRF) instead.
+Model predicting a sequence of tags can be a simple RNN with two heads - one for predicting the tags and another for predicting intent or conditional random fields (CRF) instead.
 
@@
 
-Macro average can be much lower than a weighted average, because there can be a lot of rare classes with few instances which were not predicted or predicted wrongly while there are not so many classes with a lot of instances, which are therefore predicted correctly.
+Macro average can be much lower than a weighted average (Table \ref{RNNMetrics}), because there can be a lot of rare classes with few instances which were not predicted or predicted wrongly while there are not so many classes with a lot of instances, which are therefore predicted correctly.
+
+\begin{table}
+    \centering
+    \begin{tabular}{ | c | c | c | c | c |}
+        \hline
+                        & Precision & Recall & F1-score \\
+        \hline
+        marco average   & 0.48      & 0.46   & 0.45 \\
+        \hline
+        weighted average& 0.94      & 0.95   & 0.94 \\
+        \hline
+    \end{tabular}
+    \caption{RNN Metrics}
+    \label{RNNMetrics}
+\end{table}
 

### 2022-01-18 paper.tex
inroduction

@@
 \usepackage{hyphenat}
+\usepackage{caption}
 
@@
 \pagebreak
-\section{Spoken Dialog System}
-A spoken dialogue system is a computer system that enables human computer interaction where primary input is speech.
-Examples of such systems are voice assistants, chatbots, and conversational interfaces (Siri, Alexa, Google Home, etc.).
-\par
+\section{Spoken Dialog Systems}
+A spoken dialogue system is a computer system that enables human computer interaction where primary input is speech --- voice assistants (Siri, Alexa, etc.), chatbots, and etc. \par
 
-A good spoken dialogue system has the following properties: understands the user, decides what to say back, conducts a conversation beyond simple vo\-ice commands or question answering. 
-In this overview we will be considering a goal-oriented systems (the ones which are targeted to accomplish a goal in the context of a task - to book a restaurant for ex.).
-In this case the scope of the conversation is usually limited to a specific domain - a database contating all knowledge about the topic.
-Such database is called an ontology.
-\par
-One of the straightforward and old approaches to create such a system is rule-base model.
-However such models of course require to handcraft a lot of rules, don't keep track of what has already happend in the dialog and therefore are not prone to annoying repetitive behaviour.
-And here machine learning approaches come in handy.
-\par
+A good spoken dialogue system is able to:
+\begin{itemize}
+     \item understand a user
+     \item decide what to say back
+     \item conduct a conversation beyond simple voice commands or question answering
+\end{itemize}
 
-System is usually split into modules each responsible for a specific task with well defined input and output, for each a machine learning algorithm can be used.
-Spoken language understanding (SLU), Dialogue management (DM), Spoken language generation (SLG).
-Modular approach has a downside of loosing information between the modules, for example if we have recognised speech incorrectly all following steps will be failed, however ML approaches allow to consider a list of possible inputs scored by their probability, and therefor uncertainty is propagated through the pipeline.
-\par
+Here only goal-oriented systems will be considered, where the scope of the conversation is usually limited to a specific domain with an ontology describing it.
+
+System is usually split into separate modules with well-defined input and output, in each a machine learning algo\-rithm can be used (Figure \ref{fig:SDS-main}).
+Spoken language understanding (SLU) --- for detecting and encoding users' speech, Dialogue manage\-ment (DM) --- for tracking important imformation in the dialog and choosing correct answers, Spoken language gene\-ration (SLG) --- translating answers for a user into a speech.
+Modular approach has a downside of loosing information bet\-ween the modules, however ML approaches allow to consider a list of possible inputs with probabilities, and uncertainty gets propagated through the pipeline.
+
+\begin{figure}[!h]
     \includegraphics[width=\linewidth]{SDS-main.png}
-
-Both semantic decoding and natural language generation depend on the ontology. Dialogue manager in addition needs to query the ontology to be able to provide an adequate response.
-
-\subsection{Word represeantions}
-To match user utterances with the ontology concepts we either have to either handcode all of the possible user utterances (which is completely uproductive in the constantly changing language and doesn't scale well), or learn their representation implicitely.
-For instance using a distri\-butional approach by encoding words into vectors, which can be learned statistically from the corpus of many texts.
-One can learn static representation with such statistical methods as word2vec, GloVe, fastText (uses n-grams) or contextualized neural language models such as ELMo, BERT etc.
-
-\par
-With static word respresentation one can have a meaningfull distance metric, cosine similarity for ex.: the closer the word vectors are - the more similar meaning they have.
-
-\begin{equation}
-    cos(w_i,w_j) = \frac{w_i^T \cdot w_j}{||w_i|| \cdot ||w_j||}
-\end{equation}
-
-With this approach one can get examples like: $w$("queen") = $w$("king") - $w$("man") + $w$("women")
-\par
-
-\pagebreak
-
-\section{SLU. Speech recognition}
-Speech recognition might be complicated by differences in speakers, environ\-mental noises, divestiry of vocabulary and so on.
-Uncertanties in recognition brought by these factors shoulbe be propagated through the pipeline so that information is not lost after the SLU module because of the wrong recognition.
-\begin{center}
-    \includegraphics{training-1.png}
-\end{center}
-\par 
-There is not straightforward way to convert speech into text - think for example about how to detect when user stars or ends speaking?
-\par Recognition usually doesn't assume the speech being represented in words only, but it works in the hiranhical way: utterance $\Rightarrow$ words $\Rightarrow$ subwords(pho\-nemes, triphones) $\Rightarrow$ 3 senones.
-By subwords we understand the phonemes, by triphones that one can find the in dictionary.
-3 senones represent a subwords unit - specifically its beginning, middle and end.
-
-\begin{center}
-    \includegraphics[width=\linewidth]{hierarcy-speech.png}
-\end{center}
-
-Because we have to align labels in trainging and during recognition the task is to find the most matching sequence from all possible ones we can well apply a HMM model here.
-Each phone is modeled with 3 states.
-The sequence of senones is a sequence of states for HMM.
-Each state of an HMM generates a feature vector of a frame.
-Before pure GMM were applied to solve these problems, afterwards deep learnining came into play hybrid HMM-DNN models appeared.
-There also exist end-to-end transformer models which train from sequence of phones and produce a sequence of frames.
+    \caption{}
+    \label{fig:SDS-main}
+\end{figure}
 
@@
     \includegraphics[width=\linewidth]{training-2.png}
+    \caption{}
 \end{figure}

### 2022-01-19 paper.tex
shorten first 4

@@
 
-\thispagestyle{empty}\mbox{}\pagebreak
+\thispagestyle{empty}\mbox{}
 \setcounter{page}{0}
@@
 \tableofcontents
-\pagebreak
-
-%===========================================================
-\section*{Einleitung}
-\addcontentsline{toc}{section}{Einleitung}
-
-In der vorgelegten Abschlussarbeit untersuchen wir ...
 
@@
 \section{Semantic Decoding}
-Semantic decoding translates user utterances into a set of sematic concepts (dia\-log acts).
-Dialog act consists of 2 parts: intent (dialog act type) and sematic slots with their values.
-For instance the sentence "Is there a cheap restaurant in the center?" can be translated into $inform(price=cheap,$ $area$$=center)$.
-Intent is $inform$, slot $price$ with value $cheap$ and slot $area$ with the value $center$.
+Speech recogniser module of SLU translates speech into text and outputs a list of possible user utterances along with their probabilities (Figure \ref{fig:several}).
 
-We can view semantic decoding as a classification task: having a probabi\-lity for each act type and each slot-value pair.
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.65\linewidth]{uncertanty.png}
+    \caption{}
+    \label{fig:several}
+\end{figure}
+The task of a sematic decoder module is to get the meaning from them by translating utterances into a set of sematic concepts (dia\-log acts). Dialog act consists of:
+
+\begin{itemize}
+    \item intent / dialog act type
+    \item slot-value pairs
+\end{itemize}
+
+For instance a sentence ``Is there a cheap restaurant in the center?" translates into $inform(price=cheap,$ $area$ $=center)$, where $inform$ is an intent, $price$ is a slot with value $cheap$ and $area$ is a slot with value $center$.
+
+\subsection{Models}
+One approach would be to view semantic decoding as a classification task --- having a sepa\-rate probabi\-lity for each dialog act type and each slot-value pair.
 
 \begin{figure}[h!]
-    \includegraphics[width=\linewidth]{training-2.png}
+    \centering
+    \includegraphics[width=0.8\linewidth]{training-2.png}
     \caption{}
@@
 
-Utterances for classification task can be represented as a fixed-size vector, where we preserve only information about which and how many words are there.
-To keep the word order information additionally one can use an n-gram model.
-Delexicalisation is used in advance to cluster special terms as $area$ or $food-type$ with the help of ontology.
+Utterance can be represented as a fixed-size vector, where only informa\-tion on which and how many words are there is preserved.
+To add positional information one can use an n-gram model.
+Delexicalisation clusters special terms as $area$ or $food-type$ with the help of ontology.
 Then several SVMs are trained to classify possible dialog acts (Figure \ref{fig:semantic-decoding-classification}).
@@
     \includegraphics[width=0.998\textwidth]{semantic-decoding-classification.png}
+    \caption{}
     \label{fig:semantic-decoding-classification}
@@
 
-Several possible utterances are considered (Figure \ref{several}).
-\begin{figure}[h!]
+Another approach is to view it as a seq-to-seq task ---  translating an utterance into a sequence of tags, for ex.:
+
+\begin{center}
+    ``i want to fly from Minsk to Kiev round trip tomorrow" 
+\end{center}
+
+can be translated into the sequence of tags: 
+\begin{center}
+    ["O", "O", "O", "O", "O", "B-fromloc.city\_name", "O", "B-toloc.city\_name", "B-round\_trip", "I-round\_trip", "B-departure\_date"]
+\end{center}
+\begin{figure}[!h]
     \centering
-    \includegraphics{uncertanty.png}
-    \caption{Several Utterances}
-    \label{several}
+    \includegraphics[width=0.91\linewidth]{training-2-2.png}
+    \caption{}
+    \label{fig:semantic-decoding-seq2seq}
 \end{figure}
 
-We can view this as a seq-to-seq task: translating an utterance into a sequence of tags, for ex.
-"I want to fly from Minsk to Kiev round trip tomorrow" can be translated into the sequence of tags: ["O", "O", "O", "O", "O", "B-fromloc.city\_name", "O", "B-toloc.city\_name", "B-round\_trip", "I-round\_trip", "B-departure\_date"].
-\begin{figure}[h!]
-    \includegraphics[width=\linewidth]{training-2-2.png}
-\end{figure}
+Model predicting a sequence of tags can be a simple RNN with two heads --- one for predicting the tags and another for predicting intent or conditional random fields (CRF) instead (Figure \ref{fig:semantic-decoding-seq2seq}).
 
-Model predicting a sequence of tags can be a simple RNN with two heads - one for predicting the tags and another for predicting intent or conditional random fields (CRF) instead.
+\subsection{Evaluation for RNN model}
 
-There are different ways of how we can evaluate such model.
-But first, having a sequence of tags, we can convert it to a set of slots in the following way: input = [B-food-type I\_food-type O B\_price-range O], output = [(food-type, start=0, end=1), (price-range, start=3, end=3)].
-Then having a set of true and predicted slots we can compare them on the slot level: if predicted slot was in true slots, then this is a true positive.
-If predicted slot was not in true slots, then this is a false positive.
-If true slot was not in predicted slots, then this is a false negative.
-Then from these values, we can calculate precision, recall and F1 score.
+Having predictions from RNN the easiest mectric to calculate would be a f1-score on a slot level. One first collects slots from a predicted tags sequence:
+\begin{center}
+    input = [B-food-type\quad I\_food-type\quad O\quad B\_price-range\quad O] \\
+    output = [(food-type, start=0, end=1), (price-range, start=3, end=3)]
+\end{center}
 
-On the other hand, we can calculate accuracy directly on the predicted sequence of tags, without converting it to slots.
-In this case we keep the mapping of each tag to 3 counters for TP, FP and FN.
-After that there are 2 options on how to average the metrics.
-On a macro level, we first calculate precision, recall and f1-score for each of the tags, then we average these metrics across all tags.
-On a micro level we do the weighted average, so for ex.
-to calculate precision we divide the sum of all TP by the sum of all TP+FP of all tags.
+Afterwards, having a set of true and predicted slots we can can calculate true posutives (TP), false positives (FP), false negatives (FN) on a slot level.
+Having these values one can calculate precision, recall and F1-score.
 
-Macro average can be much lower than a weighted average (Table \ref{RNNMetrics}), because there can be a lot of rare classes with few instances which were not predicted or predicted wrongly while there are not so many classes with a lot of instances, which are therefore predicted correctly.
+The other way to evaluate a model is to calculate TP, FP, FN for each tag, without converting them to slots.
+After there are two options on how to average the metrics: on a macro level (Equation \ref{eq:precision_macro}) and in a micro level (Equation \ref{eq:precision_micro}.
+Recall and F1-score are calculated analogically.
 
-\begin{table}
+\begin{equation}
+    Precision = \frac{1}{N_{slots}} \times \sum_{s \in slots}{s_{Precision}}
+    \label{eq:precision_macro}
+\end{equation}
+
+\begin{equation}
+    Precision = \frac{\sum_{s \in slots}{s_{TP}}}{\sum_{s \in slots}{s_{TP} + s_{FN}}}
+    \label{eq:precision_micro}
+\end{equation}
+
+Macro average was much lower than a micro average (Table \ref{RNNMetrics}), because there can be a lot of classes with only few instances which were not predicted or predicted wrongly while there are not so many classes with a lot of instances, which are therefore predicted correctly, therefore without weight\-ning by the number of samples (micro level) one gets a lower average.
+
+\begin{table}[!h]
     \centering
@@
                         & Precision & Recall & F1-score \\
+        \hline
+        accuracy        &           &        & 0.95 \\
         \hline
@@
     \end{tabular}
-    \caption{RNN Metrics}
+    \caption{}
     \label{RNNMetrics}
@@
 
-Let's have a look on how the tags are recognised after training. 
+Let's have a look on how the tags are recognised after training: 
 \begin{center}
@@
 
-We can fix it by adding words "with a stop in".
+We can fix it by adding words ``with a stop in".
 \begin{center}
@@
 
-There are also many instances when the state name is recognized as a city name. 
+Many instances where $state\_name$ was recognized as a $city\_name$: 
 \begin{center}
@@
 
-Arrival time can be mistaken with departing time:
+$arrive\_time$ can be mistaken with $depart\_time$:
 \begin{center}
     \scalebox{0.7}{
-    \begin{tabular}{ c c c c c}
+    \begin{tabular}{| c | c | c | c | c |}
+        \hline
         Sentence & True & Predicted & Changes & New prediction\\
+        \hline
+        \hline
         i &O&O & i & O\\
+        \hline
         would &O&O & would & O\\
+        \hline
         like  &O&O & like & O\\
+        \hline
         to &O& O & to & O\\
+        \hline
         return &O& O & fly& O\\
+        \hline
         from &O& O & around & B-depart\_time.time\_relative\\
+        \hline
         chicago &B-fromloc.city\_name& B-fromloc.city\_name & 7 & B-depart\_time.time\\
+        \hline
         around &B-depart\_time.time\_relative& B-depart\_time.time\_relative & in & O\\
+        \hline
         7 &B-depart\_time.time& B-depart\_time.time & the & O\\
+        \hline
         pm &I-depart\_time.time& I-arrive\_time.time & evening & B-depart\_time.period\_of\_day\\
+        \hline
         to &O& O & from & O\\
+        \hline
         kansas  &B-toloc.city\_name& B-toloc.city\_name & chicago & B-fromloc.city\_name\\
+        \hline
         city &I-toloc.city\_name& I-toloc.city\_name & to & O \\
+        \hline
         &&&kansas & B-toloc.city\_name\\
+        \hline
         &&& city & I-toloc.city\_name\\
+        \hline
     \end{tabular}}
@@
 
-We can fix wrongly recognized "pm" by adding - in the evening.
+We can fix wrongly recognized ``pm" by adding ``in the evening".
 %===========================================================
+\pagebreak
 \section{Dialog state tracking}
-Dialog state tracker remembers everything important that has happened in the dialog so far. Dialog management problem can be solved with the use of Markov decision process (Figure \ref{MDP}).
+First task of dialog management module is to track a dialog state.
+Dialog state tracker remembers everything important that has happened in the dialog so far.
+Tracking problem can be solved with the use of Markov decision process (Figure \ref{MDP}).
 
 \begin{figure}[ht]
-    \includegraphics[width=\linewidth]{training-state-tr-1.png}
-    \caption{MDP}
+    \centering
+    \includegraphics[width=0.8\linewidth]{training-state-tr-1.png}
+    \caption{}
     \label{MDP}
@@
 
-Process is represented as a dynamic Bayesian network (Figure \ref{Markov}) of 
+Process is represented as a dynamic Bayesian network (Figure \ref{Markov}) with 
 \begin{itemize}
-    \item Dialog states --- contain all current important information about the dialog
+    \item Dialog states --- containing current important information about the dialog
     \item Actions --- what system does
-    \item Rewards --- measures of how good is the dialog so far
+    \item Rewards --- measures of the quality of the dialog so far
 \end{itemize}
@@
       \includegraphics[width=\linewidth]{markov.png}
-      \caption{Markov}
+      \caption{}
       \label{Markov}
@@
       \includegraphics[width=\linewidth]{partial-markov.png}
-      \caption{Partial Markov}
+      \caption{}
       \label{PartialMarkov}
@@
 
-To introduce an uncertainty here, one can assume that observed utteran\-ces are not actual states, but observations, that are dependent on hidden states, then the process is called then partially observable Markov process (Figure \ref{PartialMarkov}).
+To introduce an uncertainty here, one can assume that observed utteran\-ces are dependent on hidden states, then the process is called then partially observable Markov process (Figure \ref{PartialMarkov}).
 
-Dialogue state consists of a set of distributions for the next components: 
+Dialogue state is a set of distributions for the next components: 
 \begin{itemize}
@@
 It accumulates evidence of a changing state over the course of the dialogue.
-Here we define how it tracks goals, methods, requestable slots.
+Here is defined how it tracks goals, methods, requestable slots.
 
-Define for methods and goal components $c \in V_c$ (all possible slot values) a quantity denoting that a slot or method is not present in turn $t$:
+Define for methods and goal components $c$ a quantity denoting that a slot or method is not present in turn $t$:
 
@@
 
-\noindent where $slu(v)$ and denotes a probability of given slot value $v$ for goal $c$ be\-ing a correct one.
+\noindent where $slu_c(v)$ is a probability of a given slot value $v$ for goal $c$ be\-ing a correct one, $V_c$ are all possible slot values.
 
-Recursively define probabilities $p_{c,t}(v)$ for each $v$ in $V_c \backslash \{'none'\}$ as: 
+Recursively define probabilities $p_{c,t}(v)$ for each $v \in V_c \backslash \{'none'\}$ as: 
 \begin{align}
@@
 
-For a requestable slot $s$ define 
+For requestable slots $s$: 
 \begin{align}
@@
     Using method of mathematical induction for $t=0$: \\
-    From \ref{eqn:p_c_0} since $0 \leq slu_{c,t}(v) \leq 1 \implies $
+    From (\ref{eqn:p_c_0}) since $0 \leq slu_{c,t}(v) \leq 1 \implies $
+
     \begin{equation}
@@
     \noindent Assuming that probabilities are well-defined for $t = k-1$:
+
     \begin{equation}
@@
     \end{equation}
-    Since
+
+    \noindent Since
+
     \begin{equation}
@@
     \end{equation}
-    and
-    $ slu_{c,k}(v) \geq 0 $,  $q_{c,k} \geq 0$, $p_{c,k-1}(v) \geq 0$
+
+    \noindent and $ slu_{c,k}(v) \geq 0 $,  $q_{c,k} \geq 0$, $p_{c,k-1}(v) \geq 0$
     then for $t=k$ we get
+
     \begin{equation}
@@
     \end{equation}
-    As $p_{c, t}('none') \leq 1$ by definition (\ref{eqn:p_c_t_none}), we only need to prove it being $\geq 0$.\\
 
-    \noindent Again by method of mathematical induction let's assume that for $t = k-1$ $p_{c, k-1}('none') \geq 0$ then for $t = k$
+    \noindent As $p_{c, t}('none') \leq 1$ by definition (\ref{eqn:p_c_t_none}), we only need to prove it being $\geq 0$.\\
+
+    \noindent Again by method of mathematical induction assuming that for $t = k-1$ $p_{c, k-1}('none') \geq 0$ then for $t = k$:
 
@@
 
-    For requestable slots.\\
+    For requestable slots. \\
 
     Assuming that probabilities are well-defined for $t = k-1$:
+
     \begin{equation}
@@
 
-    Since
+    \noindent Since
 
@@
     \end{equation}
+
+    \noindent Therefore
 
@@
 \subsection{Quality Evaluation}
-To evaluate a quality of a belief state tracker one can use a fraction of turns where a top dialogue state hypothesis was correct (accuracy) or use the following equa\-tion (\ref{l2}) where $p$ is a hypothesised distribution, $i$ is an index of a true label.
+For quality evaluation one can use either a fraction of turns where a top dialogue state hypothesis was correct (accuracy) or alternatively:
 \begin{equation}
@@
 
-In comparison to the simplest tracker focus tracker performs much better for informable slots.
-This is an expected behaviour as informable slot may change over time, and probabi\-lity of the slot being informable will also change.
-However for the requestable slot focus tracker doesn't show a better performance as usually the requestable slots should not change during the dialog, they are requested once and after that removed from the set of the requestable slots, after that the user can decide to request another one.
-The same applies for methods, performance there is similar to the baseline tracker as user only once "sets" by which method he wants to achieve the goal.
+\noindent
+where $p$ is a hypothesised distribution, $i$ is an index of a true label.
 
-One can see the results for both trackers in Table \ref{focus_tracker_results} and Table \ref{baseline_tracker_results}.
+Focus tracker in comparison to the simplest tracker performs much better for goals (sometimes called informable slots) as they may change over time, for.
+ex.
+user may ask for another pricerange or another restaurant area.
+However for requestable slots its performance is not better as these slots usually should not change during the dialog, they are requested once and after that removed from the set of the requestable slots, after that the user can decide to request another one.
+Similar for methods as the user usually "sets" by which method he will achieve the goal once (Table \ref{focus_tracker_results} and Table \ref{baseline_tracker_results}).
+
 \begin{table}[ht]
@@
 \end{table}
-
 \begin{table}[ht]
@@
 \subsection{Interesting Question}
-What happens to a probability when user gives an exact same utterance in
-the first two turns of the dialog?
+Assume user gives an exact same utterance in the first two turns of the dialog, what happens to a probability?
 
-In this case $slu(v) = slu_{c, 1}(v) = slu_{c,2}(v) \Rightarrow q_{c,1} = q_{c, 2} = q$ for $\forall c \forall v \in V_c $
+$slu(v) = slu_{c, 1}(v) = slu_{c,2}(v) \Rightarrow q_{c,1} = q_{c, 2} = q$ for $\forall c \forall v \in V_c $
 
+\noindent Then
 \begin{align*}
@@
 
-$\Rightarrow$ probability of value $v$  for slot $c$ will grow $p_{c,2}(v) \geq p_{c,1}(v) \forall v \in V_c $
+\noindent $\Rightarrow$ probability of value $v$  for slot $c$ will grow $p_{c,2}(v) \geq p_{c,1}(v) \forall v \in V_c $
 
@@
 
-\noindent and the probability of value $none$ decreases.
-This makes an intuitie sense, if a user repeats a sentence an agen should increase a probability of the slots mentioned by him and descrease a probability of $none$ values as user definitely wants to pass to the agent some important information.
+\noindent Probability of value $none$ decreases.
+This is intuitive, with a sentence repetition an agent should increase a probability of the slots mentioned by user and descrease a probability of $none$ values as user definitely insist on some important information.
 
 \subsection{Another update rule}
-Let's define another update rule:
+Define an update rule:
 \begin{equation}
@@
 
-Since (\ref{eqn:q_c_0_1}) means that all probabilities were 0 for any $v$ apart from $none$, then $p_{c,0}(v) = 0$ for all $v \in V_c$. Then: 
+Since (\ref{eqn:q_c_0_1}) means that $p_c(v) = 0  \quad \forall v \in V_c\{'none'\} $, then: 
 \begin{align*}
@@
 
-As the slot was not in the first turn and then appeared in the second turn with probability 0.5, then it is okay to assume that the the probability of it might be not 0.5 (as it would be in previous update rule) but 0.25.
+As the slot was not in the first turn and then appeared in the second turn with probability 0.5, it is reasonable to assume that its probability might be not 0.5 (by previous update rule) but 0.25.
 
-In general, such update rule puts less weight on the new probabilities from current turn and more weight on the previous turn, which is also a totally reasonable update.
+In general, such update rule puts less weight on the new probabilities from current turn and more weight on the previous turn, which is also a reasonable update.
 
@@
 
-\section{Dialog state tracking as a reinforcement learning problem}
-Apart from tracking a state of a dialog, the agent must learn to perform the correct action.
-This is can be viewed as a reinforcement learning problem, where the agent is rewarded for performing correct actions and might be penalised for the incorrect ones.
-Agent owns a policy (its behaviour) that it has to optimize, rewards from the environment, value function (prediction of the future reward given current state $b$) and a model of the environment.
+\section{Dialog optimisation as a reinforcement learning problem}
+Apart from tracking a state of a dialog, the agent must learn to perform correct actions.
+This can be viewed as a reinforcement learning problem, where the agent is rewarded for performing correct actions and penalised for incorrect ones.
+Agent has a policy (its behaviour) that should be optimized, rewards from the environment, $V$-function (prediction of the future reward given current state $b$) and a model of environment.
 
-One also defines a Q-function, which determines how good it is to take action $a$ in belief state $b$ and then follow a policy $\pi$.
-Having an optimal Q-function one chooses an action that maximises its value, this will be an optimal action.
+$Q$-function is also defined to determines how good is it to take action $a$ in belief state $b$ and then follow a policy $\pi$. Having an optimal $Q$-function one can choose an optimal action.
 
@@
 
-\subsection{Why do to define a reward?}
+\subsection{Rewards}
+Reward is a measure of success of the policy for the dialog, without reward one cannot learn an optimal policy.
+By optimizing reward value one opti\-mizes the policy.
+It is a metric for the algorithm's success.
 
-Reward is a measure of success of the policy for the dialog, if we don’t have a reward, we cannot learn and define an optimal policy.
-With it we can define how we optimize the policy by optimizing the value of the reward.
-It is basically a metric for the algorithm's success.
+\subsection{Allowed actions for execution}
 
-\subsection{What actions can system execute?}
-
-Interesting trick for optimizing a training process is to explicitely define which actions are allowed for execution.
-An agent should be not allowed to execute those actions which are clearly impossible.
-For example, after user requests a restaurant's pnone number an agent should not be allowed to execute an action “say hello”.
-In order to implement that we can mask out with 0 those actions that are not allowed to be executed.
+For optimization one can restrict some actions to be executed as an agent should be not allowed to execute those actions which are clearly implausible.
+For example, after user requests a restaurant's address an agent should not say ``hello”.
+For that one can mask out with 0 those actions that are not allowed to be executed.
 
@@
 
-Monte Carlo control algorithm is a method for finding the optimal policy for a reinforcement problem.
+Monte-Carlo control algorithm finds an optimal policy for a reinforcement learning problem.
 
-It is a tabular approach, which means that the belief state/summary space is discretised into a grid.
-In this algorithm the policy and the Q function are first defined randomly.
-Then we iterate over the dialogs with this polilcy.
-Throught this iteration we calculate the reward until the end of the dialog.
-For updating the state we choose the nearest state from the table and update the Q value and afterwards the policy with it, then in the next turn we will use an already updated policy.
+It is a tabular approach, meaning that the belief state space is discretised into a grid.
+Following a random policy and $Q$-function one generates a dialog and rewards, which are iteratively used to update a $Q$-function and a policy afterwards.
+For updates the nearest state from the table is chosen (Figure \ref{fig:MCC}).
 
-\begin{center}
+\begin{figure}[!h]
+    \centering
     \includegraphics[width=10cm]{MCC.png}
-\end{center}
+    \caption{}
+    \label{fig:MCC}
+\end{figure}
 

### 2022-01-19 paper.tex
shorten 2025

@@
 %===========================================================
-\section{Deep Q-Network}
+\section{Deep Q-Learning}
 
 %===========================================================
-\subsection{Q-Learning}
+\subsection{Q-Network}
 Instead of keeping Q-values in a table, one can approximate a Q-function with a neural network (Q-network).
-It is a neural network that takes a state of a dialog as input and outputs Q-values for each possible action.
+It takes a state of a dialog as input and outputs Q-values for each possible action.
 An optimal Q-network obeys the following Bellman optimality equation:
@@
 
-where $\gamma$ is a discount factor,  $r$ is an immediate reward and $\max_a Q(s',a)$ is the maximum Q-value of the next state. One uses Bellman equation to update the Q-function iteratively:
+where $\gamma$ is a discount factor,  $r$ is an immediate reward and $\max_a Q(s',a)$ is the maximum Q-value of the next state. Bellman equation is used to update the Q-function iteratively:
 
@@
 
-With iterative updates it will converge to an optimal Q-function.
+\subsection{Results}
 
-In comparison to Monte-Carlo this approach this one is more efficient as it does not keep track of all the past states.
-Instead it only remembers the current one.
-This happens because only Q-value of the current state is used to calculate a Q-value of the next state.
-Monte-Carlo reaches the maximum success rate of TODO 76\% close to the 2000 dialogs, whereas Q-learning reaches the maximum success rate of 96\% after 2000 dialogs and has already 94\% after ~400 dialogs.
-The reward has almost doubled as well.
+In comparison to Monte-Carlo this approach is more efficient as it does not keep track of all the past states and remembers only the current one, because only Q-value of the current state is used for the next step update.
+Q-Network trains faster and better then a Monte-Carlo approach (TODO 96\% vs.
+76\% Figures \ref{fig:ex}, \ref{fig:mc}).
 
@@
 For experience replay one stores tuples ${s, a, r, s^\prime}$ in the buffer, where $s^\prime$ is the following dialog state and samples batches from the it.
-Delayed targets allow to stabilize the network as we the target network’s parameters are not trained, but they are periodically synchronized with the parameters of the main Q-network.
-The loss funciton here is the difference between Q-value of the current state + current reward vs.
-Q-value of the next state:
+Delayed targets allow to stabilize the network as we the target network’s parameters are not trained, but periodically synchronized with the main Q-network.
+Loss function:
 
@@
 
-where $Q_{target}(r, s^\prime) = r + \gamma max_{a}Q_{target}(s^\prime,a)$
+where
 
-if $s^\prime$ is a terminal state, then $Q(s^\prime, a) = Q_{target}(s^\prime, a) = 0$ for every $a$.
+\[Q_{target}(r, s^\prime) = r + \gamma max_{a}Q_{target}(s^\prime,a)\]
 
-Additionally, we have trained a Q-network without experience replay.
-In this case reward was almost 3 times lower TODO (11.23 and 4.5) as well as the success rate.
-The variance has also increased and therefore the network became more unstable.
-One can see the results in Table TODO.
-
-\begin{center}
+Let's train a Q-network without experience replay.
+In this case reward was almost 3 times lower TODO (11.23 and 4.5 Figure \ref{fig:no-ex}, Table \ref{tab:replay-no-replay}).
+With increased variance the $Q$-network became more unstable.
+\begin{table}[!h]
+    \centering
     \begin{tabular}{||c c c c||} 
@@
     \end{tabular}
-\end{center}
-
-The same informations is presented also in the plots.
+    \caption{}
+    \label{tab:replay-no-replay}
+\end{table}
 
@@
       \caption{Experience Replay}
+      \label{fig:ex}
       \endminipage\hfill
@@
       \caption{No Experience Replay}
+      \label{fig:no-ex}
       \endminipage\hfill
@@
       \caption{Monte-Carlo}
+      \label{fig:mc}
     \endminipage
@@
 
-One can see that system sometimes is caught in the loop and keeps asking the same question over and over again.
-The reason might be that semantic decoder didn't work well and haven't recognised an imput correctly, so the systems continue to request more information.
+\subsection{Chatting with a Q-network}
+Here are presented examples of dialogs conducted with a Q-network.
+System sometimes gets caught in the loop and keeps asking the same question.
+The reason might be that semantic decoder didn't work well and haven't recog\-nised an input correctly, so the system continues to request more information.
 TODO get example
 
-There was no instances in the dataset where the user would say "I don't care" regarding the price range, and we can see an example of how the dialog is ruined when we encounter such a case, compare dialog 1 vs dialog 2.
-TODO dialog number Usually such infinite loops may happen due to bad recognition of speech, however here the problem is hidden probably in the state tracker.
+There was no instances in the dataset where the user wouldn't care which price range is there, and the dialog is ruined when this case is encountered (Dialog \ref{diag:with-price} vs Dialog \ref{diag:without-price}).
+This might happen due to insufficient semantic decoder knowledge and not enough of such samples in the database.
 
@@
       \includegraphics[width=\linewidth]{with_price.png}
-      \caption{With price}
+      \caption{}
+      \label{diag:with-price}
     \endminipage\hfill
@@
       \includegraphics[width=\linewidth]{without_price.png}
-      \caption{Without price}
+      \caption{}
+      \label{diag:without-price}
     \endminipage
@@
 
-Another noticable thing is that choice of value for one slot sometimes influences the dialog drastically even though the slot is not that relevant.
-For example, here in the dialog 3 we choose specific price in comparison with dialog 4 TODO check number where we are agreed to any price and that leads suddenly to an infinite loop.
-This might happen due to insufficient semantic decoder knowledge and not enough of such samples in the database.
-\begin{figure}[!htb]
-    \minipage{0.49\textwidth}
-      \includegraphics[width=\linewidth]{chinese_works.png}
-      \caption{Working sample}
-    \endminipage\hfill
-    \minipage{0.49\textwidth}%
-      \includegraphics[width=\linewidth]{chinese_no_work.png}
-      \caption{Not working sample}
-    \endminipage
-\end{figure}
-
-Examples with negations are performing poorly as well which is also most probably the failure of the semantic decoder (see dialog 5 TODO change).
+Examples with negations are performing poorly, most probably the failure of the semantic decoder not recognizing negation and not enough of negative samples in the database (Dialog \ref{diag:negation}).
 
 \begin{figure}
+    \centering
     \includegraphics[width=5cm]{chinese\ negation.png}
+    \caption{}
+    \label{diag:negation}
 \end{figure}
@@
 \pagebreak
+
+\section{Conclusion}
+
 \begin{thebibliography}{ccccc}

### 2022-01-19 paper.tex
shorten 1935

@@
 \section{Dialog state tracking}
-First task of dialog management module is to track a dialog state.
-Dialog state tracker remembers everything important that has happened in the dialog so far.
-Tracking problem can be solved with the use of Markov decision process (Figure \ref{MDP}).
+Dialog state tracker os a part of dialog management module and remembers everything important that has happened in the dialog so far.
+Markov decision process (Figure \ref{MDP}) might be applied for tracking problem.
 
@@
 
-Process is represented as a dynamic Bayesian network (Figure \ref{Markov}) with 
+Process is represented as a dynamic Bayesian network (Figure \ref{Markov}) of: 
 \begin{itemize}
-    \item Dialog states --- containing current important information about the dialog
+    \item Dialog states --- current important information
     \item Actions --- what system does
-    \item Rewards --- measures of the quality of the dialog so far
+    \item Rewards --- quality of the dialog so far
 \end{itemize}
@@
 
-To introduce an uncertainty here, one can assume that observed utteran\-ces are dependent on hidden states, then the process is called then partially observable Markov process (Figure \ref{PartialMarkov}).
+One can assume that observed utteran\-ces are dependent on hidden states, then the process is called then partially observable Markov process (Figure \ref{PartialMarkov}).
 
@@
 \begin{itemize}
-    \item Goal --- for each slot a distribution over possible values
-    \item Method --- a distribution over possible methods
-    \item Requested slots --- for each requestable slot a binary distribution
+    \item Goal --- distribution over possible values (for each slot)
+    \item Method --- distribution over possible methods
+    \item Requested slots --- binary distribution (for each slot)
 \end{itemize}
@@
 
-The simplest tracker assigns a single hypothesis for each slot at each turn, a more advanced option is a focus tracker.
-It accumulates evidence of a changing state over the course of the dialogue.
-Here is defined how it tracks goals, methods, requestable slots.
-
-Define for methods and goal components $c$ a quantity denoting that a slot or method is not present in turn $t$:
+The simplest tracker assigns a single hypothesis for each slot at each turn, whereas a focus tracker accumulates evidence of a changing state over the course of the dialogue.
+It works in the following way: define for methods, goal components $c$ a quantity denoting that a slot / method is not present in turn $t$:
 
@@
 
-\noindent where $slu_c(v)$ is a probability of a given slot value $v$ for goal $c$ be\-ing a correct one, $V_c$ are all possible slot values.
+\noindent where $slu_c(v)$ --- a probability of a given slot value $v$ for goal $c$ be\-ing a correct one, $V_c$ --- all possible slot values.
 
-Recursively define probabilities $p_{c,t}(v)$ for each $v \in V_c \backslash \{'none'\}$ as: 
+Define $p_{c,t}(v)$ for each $v \in V_c \backslash \{'none'\}$ as: 
 \begin{align}
@@
 
-Now let's prove that probabilities are well-defined, in other words that $0 \leq p_{c,t}(v) \leq 1$ $\forall v \in V_c, \forall t, \forall c$
+Let's prove that probabilities are well-defined, meaning that $0 \leq p_{c,t}(v) \leq 1$ $\forall v \in V_c, \forall t, \forall c$
 
@@
 
-    \noindent Assuming that probabilities are well-defined for $t = k-1$:
+    \noindent Assuming that for $t = k-1$:
 
@@
     \noindent and $ slu_{c,k}(v) \geq 0 $,  $q_{c,k} \geq 0$, $p_{c,k-1}(v) \geq 0$
-    then for $t=k$ we get
+    then for $t=k$:
 
@@
 
-    \noindent As $p_{c, t}('none') \leq 1$ by definition (\ref{eqn:p_c_t_none}), we only need to prove it being $\geq 0$.\\
+    \noindent As $p_{c, t}('none') \leq 1$ from (\ref{eqn:p_c_t_none}), let's prove $p_{c, t}('none') \geq 0$.\\
 
-    \noindent Again by method of mathematical induction assuming that for $t = k-1$ $p_{c, k-1}('none') \geq 0$ then for $t = k$:
+    \noindent By method of mathematical induction assuming that for $t = k-1$ $p_{c, k-1}('none') \geq 0$ then for $t = k$:
 
@@
 
-    Assuming that probabilities are well-defined for $t = k-1$:
+    Assuming that for $t = k-1$:
 
@@
 ex.
-user may ask for another pricerange or another restaurant area.
-However for requestable slots its performance is not better as these slots usually should not change during the dialog, they are requested once and after that removed from the set of the requestable slots, after that the user can decide to request another one.
+user may ask for another pricerange, location.
+However for request\-able slots its performance is not better as these slots usually should not change during the dialog, they are requested once, removed from the set, afterwards user can request next one.
 Similar for methods as the user usually "sets" by which method he will achieve the goal once (Table \ref{focus_tracker_results} and Table \ref{baseline_tracker_results}).
@@
 
-\noindent $\Rightarrow$ probability of value $v$  for slot $c$ will grow $p_{c,2}(v) \geq p_{c,1}(v) \forall v \in V_c $
+\noindent $\Rightarrow$ probability of value $v$  for slot $c$ increases $p_{c,2}(v) \geq p_{c,1}(v) \forall v \in V_c $
 
@@
 \noindent Probability of value $none$ decreases.
-This is intuitive, with a sentence repetition an agent should increase a probability of the slots mentioned by user and descrease a probability of $none$ values as user definitely insist on some important information.
+This is intuitive --- by a repetition user insists on some important information and agent becomes more confident in mentioned values and less in value $none$.
 
@@
 \end{equation}
-Assume we talked to the user for two turns and for an informable slot $c$ it holds that 
+Assume two turns of dialog happened and that for an informable slot $c$:
 \begin{equation}
@@
 
-Since (\ref{eqn:q_c_0_1}) means that $p_c(v) = 0  \quad \forall v \in V_c\{'none'\} $, then: 
+Since (\ref{eqn:q_c_0_1}) $\Rightarrow$ $p_c(v) = 0  \quad \forall v \in V_c\{'none'\} $, then: 
 \begin{align*}
@@
 
-In general, such update rule puts less weight on the new probabilities from current turn and more weight on the previous turn, which is also a reasonable update.
+In general, such update rule puts less weight on the new probabilities from current turn and more weight on the previous turn, which is also reasonable.
 
@@
 
-\section{Dialog optimisation as a reinforcement learning problem}
+\section{Dialog optimisation as a reinforcement learn\-ing problem}
 Apart from tracking a state of a dialog, the agent must learn to perform correct actions.
@@
 
+\pagebreak
+
 \section{Information Gain}
@@
 This reward is called information gain and it measures how much additional information the system gets about user needs after one turn.
-Let action be of form $\{request, confirm,$ $select\}\_\xi$.
-Let $p_{\xi}$ and ${p_{\xi}}^\prime$ be belief distributions over possible values for $\xi$ in current and next belief state.
-Then the informa\-tion gain is the difference between these distributions with Jensen-Shannon divergence as a distance metric:
+Let action be of form $\{request, confirm,$ $select\}\_\xi$, $p_{\xi}$ and ${p_{\xi}}^\prime$ --- belief distributions over possible values of $\xi$ in current and next belief state.
+Informa\-tion gain then is the distance between these distributions measured by with Jensen-Shannon di\-vergence distance:
 
@@
 
-When the distance is big enough (higher than a threshold) we add the information gain to a reward additionally.
-Reward will be added only when the action was of the form ${request, confirm, select}$.
-Intuitively it also doesn't make sense to calculate information gain for actions like $bye, inform$ as they do not provide any additonal information on user needs and therefore do not lead to the accomplishment of the dialog goal.
-It is very beneficial to use intrinsic rewards in addition as they encourage system to get as much useful information from users as possible, by requesting more information from them, selecting the most relevant one and confirming the choices.
+When distance is big enough (higher than a threshold) we add infor\-mation gain to the reward, but only when the action was of the form $\{request,$ $ confirm, select\}$.
+Intuitively it also doesn't make sense to calculate infor\-mation gain for actions like $bye, inform$ as they do not provide any additonal infor\-mation on user needs and therefore do not lead to the accomplishment of the dialog goal.
+It is beneficial to use intrinsic rewards in addition as they encourage system to get more useful information from users by requesting more from them, selecting the most relevant one information and confirming the choices.
 
 Here one can see the difference in training with information gain and without it.
-Information gain improves the speed of training in the beginning and slightly improves the overall performance from $96\%$  to $98.4\%$ (see Figure 7, 8).
+Information gain improves the speed of training and slightly improves an overall performance from $96\%$  to $98.4\%$ (see Figure 7, 8).
 

### 2022-01-19 paper.tex
conclusion

@@
 
-\textbf{\LARGE Titel der Bachelorarbeit}
+\textbf{\LARGE Spoken Dialog Systems}
 %bei langen Titeln, die mehrere Zeilen benoetigen:
@@
 \bigskip\bigskip
-\textbf{Bachelorarbeit}
+\textbf{Report}
 %bei Masterarbeit:
@@
 \bigskip\bigskip\bigskip
-Vorgelegt von
+Written by
 
 \bigskip
-\textbf{Vorname Nachname}
-
-\bigskip
-aus Geburtsort
+\textbf{Hanna Pankova}
 
 \vfill
-Angefertigt am\\
-Mathematischen Institut\\ 
-der Mathematisch-Naturwissenschaftlichen Fakult\"at\\ 
-der Heinrich-Heine-Universit\"at D\"usseldorf
+Faculty of Mathematics and Natural Sciences\\ 
+Heinrich Heine University \\
+D\"usseldorf
 
@@
 %Abgabedatum:
-tt.\ Monat jjjj
+01 February 2022
 
 \bigskip
-Betreuer: Prof.\ Dr.\ Vorname Nachname
+Lecturer: Prof.\ Dr.\ Milica Ga\v{s}i\'{c}
 %Wird der Schwerpunkt der Abschlussarbeit im Anwendungsfach gewaehlt:
@@
 \section{Conclusion}
+This report gives a brief introduction to the construction of spoken dialog systems.
+It contains research on different approaches and algorithms used for end-to-end spoken dialog system, that are build via modular approach.
+Such system was built with the use of different deep learning approaches, all approaches for each of the modules were studied and compared, several dialogs were conducted with the system.
+The results of the study were described and presented.
 
+\pagebreak
 \begin{thebibliography}{ccccc}
@@
 \end{thebibliography}
-
-%===========================================================
-\pagebreak\noindent
-\textbf{\LARGE Erkl\"arung}
-\addcontentsline{toc}{section}{Erkl\"arung}
-
-\bigskip\bigskip
-\noindent 
-Hiermit versichere ich, dass ich die   Bachelorarbeit selbst\"andig verfasst und keine
-anderen als die angegebenen Quellen und Hilfsmittel benutzt habe.
-
-\bigskip
-\noindent
-D\"usseldorf, den tt.\ Monat jjjj
-
-\bigskip\bigskip\bigskip
-\noindent
-(Vorname Nachname)
-
 \end{document}

### 2022-01-20 paper.tex
.

@@
 \usepackage{amsfonts}
-\usepackage{babel}
+\usepackage[english]{babel}
 \usepackage{graphicx}
@@
      \item understand a user
-     \item decide what to say back
+     \item decide what to answer
      \item conduct a conversation beyond simple voice commands or question answering
@@
 \section{Semantic Decoding}
-Speech recogniser module of SLU translates speech into text and outputs a list of possible user utterances along with their probabilities (Figure \ref{fig:several}).
+Speech recogniser module of SLU translates speech into text and outputs a list of possible user utterances along with their probabilities using a hidden markov model (HMM) (Figure \ref{fig:several}).
+REFERENCE Gales, M. and Young, S. (2007).
+The application of hidden markov models in speech recognition.
+https://mi.eng.cam.ac.uk/~mjfg/mjfg_NOW.pdf
 
@@
 \end{center}
+REFERENCE https://static.googleusercontent.com/media/research.google.com/en//pubs/archive/44628.pdf
+Mesnil, G., Dauphin, Y., Yao, K., Bengio, Y., Deng, L., Hakkani-Tur, D., He, X., Heck, L., Tur, G., Yu, D., and Zweig, G. (2015).
+Using recurrent neural networks for slot filling in spoken language understanding.
+
 \begin{figure}[!h]
@@
 \end{figure}
+REFERENCE Tur, G., Deoras, A., and Hakkani-Tur, D. (2013).
+Semantic parsing using word confusion networks with conditional random fields.
 
@@
 Such system was built with the use of different deep learning approaches, all approaches for each of the modules were studied and compared, several dialogs were conducted with the system.
-The results of the study were described and presented.
+The results of the study were described and presente.
 
 \pagebreak
+
 \begin{thebibliography}{ccccc}
@@
 \end{thebibliography}
+
 \end{document}

### 2022-01-20 paper.tex
some references

--- 2022-01-19 paper.tex
@@
 \usepackage{caption}
+\usepackage[numbers]{natbib}
 
@@
 \section{Semantic Decoding}
-Speech recogniser module of SLU translates speech into text and outputs a list of possible user utterances along with their probabilities (Figure \ref{fig:several}).
+Speech recogniser module of SLU translates speech into text and outputs a list of possible user utterances along with their probabilities using a hidden markov model (HMM)\cite{HMM} (Figure \ref{fig:several}).
 
@@
 \end{center}
+Using recurrent neural networks for slot filling in spoken language understanding.
+
 \begin{figure}[!h]
@@
 
-Model predicting a sequence of tags can be a simple RNN with two heads --- one for predicting the tags and another for predicting intent or conditional random fields (CRF) instead (Figure \ref{fig:semantic-decoding-seq2seq}).
+Semantic parsing using word confusion networks with conditional random fields.
+
+Model predicting a sequence of tags can be a simple RNN \cite{SemanticFilling} with two heads --- one for predicting the tags and another for predicting intent or conditional random fields (CRF) \cite{CRF} instead (Figure \ref{fig:semantic-decoding-seq2seq}).
 
@@
 \pagebreak
-\begin{thebibliography}{ccccc}
-\addcontentsline{toc}{section}{Literatur}
 
-%So kann eine eine Monographie zitieren:
-\bibitem{Arbarello et al 1985}
-E.\ Arbarello, M.\ Cornalba, P.\ Griffiths, J.\ Harris:
-Geometry of algebraic curves. I. 
-Springer, New York, 1985.
+\bibliography{refs} 
+\bibliographystyle{plainnat}
 
-%So kann man eine Originalarbeit zitieren:
-\bibitem{Shepherd-Barron 1997}
-N.\ Shepherd-Barron:
-Fano threefolds in positive characteristic.
-Compositio Math.\  105  (1997),  237--265.
-
-\end{thebibliography}
 \end{document}

### 2022-01-20 paper.tex
finita la comedia

@@
 \usepackage[numbers]{natbib}
+\usepackage[printonlyused,withpage]{acronym}
 
@@
 \pagebreak
+
 \section{Spoken Dialog Systems}
-A spoken dialogue system is a computer system that enables human computer interaction where primary input is speech --- voice assistants (Siri, Alexa, etc.), chatbots, and etc. \par
+A spoken dialogue system (SDS) is a computer system that enables human computer interaction where primary input is speech --- voice assistants (Siri, Alexa, etc.), chatbots, and etc. \par
 
-A good spoken dialogue system is able to:
+A good SDS is able to:
 \begin{itemize}
@@
 
-Here only goal-oriented systems will be considered, where the scope of the conversation is usually limited to a specific domain with an ontology describing it.
+Here will be considered only goal-oriented SDSs, where the scope of the conversation is usually limited to a specific domain described by an ontology.
 
-System is usually split into separate modules with well-defined input and output, in each a machine learning algo\-rithm can be used (Figure \ref{fig:SDS-main}).
-Spoken language understanding (SLU) --- for detecting and encoding users' speech, Dialogue manage\-ment (DM) --- for tracking important imformation in the dialog and choosing correct answers, Spoken language gene\-ration (SLG) --- translating answers for a user into a speech.
-Modular approach has a downside of loosing information bet\-ween the modules, however ML approaches allow to consider a list of possible inputs with probabilities, and uncertainty gets propagated through the pipeline.
+System is usually split into separate modules \cite{Structure} with well-defined input and output.
+In each can be used a machine learning (ML) algo\-rithm (Figure \ref{fig:SDS-main}).
+Spoken language understanding (SLU) --- for detecting and encoding users' speech, Dialogue manage\-ment (DM) --- for tracking important imformation in the dialog and choosing correct answers, Spoken Language gene\-ration (SLG) --- translating answers into speech.
+However, such modular approach has a downside of loosing information bet\-ween the modules, luckily since ML approaches allow to consider a list of possible inputs with probabilities, and uncertainties get propagated through the pipeline.
 
 \begin{figure}[!h]
+    \centering
     \includegraphics[width=\linewidth]{SDS-main.png}
-    \caption{}
+    \caption{Modular spoken dialogue system}
     \label{fig:SDS-main}
@@
     \includegraphics[width=0.65\linewidth]{uncertanty.png}
-    \caption{}
+    \caption{List of possible utterances from speech recogniser}
     \label{fig:several}
 \end{figure}
-The task of a sematic decoder module is to get the meaning from them by translating utterances into a set of sematic concepts (dia\-log acts). Dialog act consists of:
+The task of a sematic decoder is to get the meaning from the utterances by translating them into a set of sematic concepts (dia\-log acts). Dialog act consists of:
 
@@
     \includegraphics[width=0.8\linewidth]{training-2.png}
-    \caption{}
+    \caption{Semantic decoding as a classification task}
 \end{figure}
@@
     \includegraphics[width=0.998\textwidth]{semantic-decoding-classification.png}
-    \caption{}
+    \caption{Classification using SVM}
     \label{fig:semantic-decoding-classification}
@@
 
-can be translated into the sequence of tags: 
+can be translated into the sequence of tags \cite{IOB}: 
 \begin{center}
@@
 \end{center}
-Using recurrent neural networks for slot filling in spoken language understanding.
 
+Such slot filling in SLU could be done by using recurrent neural networks (RNN) or word confusion networks with conditional random fields (CRF) \cite{CRF}.
 \begin{figure}[!h]
@@
     \includegraphics[width=0.91\linewidth]{training-2-2.png}
-    \caption{}
+    \caption{Classification using RNN or CRF}
     \label{fig:semantic-decoding-seq2seq}
@@
 
-Semantic parsing using word confusion networks with conditional random fields.
+RNN \cite{SemanticFilling} can be a simple network with two heads --- one for predicting the tags and another for predicting intent (Figure \ref{fig:semantic-decoding-seq2seq}).
 
-Model predicting a sequence of tags can be a simple RNN \cite{SemanticFilling} with two heads --- one for predicting the tags and another for predicting intent or conditional random fields (CRF) \cite{CRF} instead (Figure \ref{fig:semantic-decoding-seq2seq}).
+\subsection{Evaluation of RNN Models}
 
-\subsection{Evaluation for RNN model}
-
-Having predictions from RNN the easiest mectric to calculate would be a f1-score on a slot level. One first collects slots from a predicted tags sequence:
+The easiest metric for RNN predictions evaluation would be a F1-score on a slot level. One first collects slots from a predicted tags sequence:
 \begin{center}
     input = [B-food-type\quad I\_food-type\quad O\quad B\_price-range\quad O] \\
+    translates into \\
     output = [(food-type, start=0, end=1), (price-range, start=3, end=3)]
@@
 The other way to evaluate a model is to calculate TP, FP, FN for each tag, without converting them to slots.
-After there are two options on how to average the metrics: on a macro level (Equation \ref{eq:precision_macro}) and in a micro level (Equation \ref{eq:precision_micro}.
+After there are two options on how to average the metrics: on a macro level (Equation \ref{eq:precision_macro}) and in a micro level (Equation \ref{eq:precision_micro}).
 Recall and F1-score are calculated analogically.
@@
 
-Macro average was much lower than a micro average (Table \ref{RNNMetrics}), because there can be a lot of classes with only few instances which were not predicted or predicted wrongly while there are not so many classes with a lot of instances, which are therefore predicted correctly, therefore without weight\-ning by the number of samples (micro level) one gets a lower average.
+Macro average was much lower than a micro average (Table \ref{RNNMetrics}), because there can be a lot of classes with only few instances which were not predicted or predicted wrongly, while there are not so many classes with a lot of instances, which were predicted correctly, therefore without weight\-ning by the number of samples (micro level) one gets a lower average.
 
@@
     \end{tabular}
-    \caption{}
+    \caption{RNN training results}
     \label{RNNMetrics}
@@
 
-Let's have a look on how the tags are recognised after training: 
+This is how tags were recognised after training: 
 \begin{center}
@@
 
-We can fix it by adding words ``with a stop in".
+This prediction was a bit off, one can fix it by adding words ``with a stop in".
 \begin{center}
@@
 
-We can fix wrongly recognized ``pm" by adding ``in the evening".
+One can fix wrongly recognized ``pm" by adding ``in the evening".
 %===========================================================
 \pagebreak
-\section{Dialog state tracking}
-Dialog state tracker os a part of dialog management module and remembers everything important that has happened in the dialog so far.
-Markov decision process (Figure \ref{MDP}) might be applied for tracking problem.
+\section{Dialog State Tracking}
+Dialog state tracker (DST) is a part of DM module and remembers everything important that has happened in the dialog so far.
+Markov decision process (MDP) (Figure \ref{MDP}) might be applied for DST problem.
 
@@
     \includegraphics[width=0.8\linewidth]{training-state-tr-1.png}
-    \caption{}
+    \caption{DST as a Markov decision process}
     \label{MDP}
@@
 
-One can assume that observed utteran\-ces are dependent on hidden states, then the process is called then partially observable Markov process (Figure \ref{PartialMarkov}).
+One can assume that observed utteran\-ces are dependent on hidden states, then the process is called then partially observable MDP (Figure \ref{PartialMarkov}).
 
-Dialogue state is a set of distributions for the next components: 
+Dialogue state is a set of distributions for the next components \cite{DST}: 
 \begin{itemize}
@@
 
-The simplest tracker assigns a single hypothesis for each slot at each turn, whereas a focus tracker accumulates evidence of a changing state over the course of the dialogue.
+The simplest tracker assigns a single hypothesis for each slot at each turn \cite{DST}, whereas a focus tracker accumulates evidence of a changing state over the course of the dialogue.
 It works in the following way: define for methods, goal components $c$ a quantity denoting that a slot / method is not present in turn $t$:
@@
 \subsection{Quality Evaluation}
-For quality evaluation one can use either a fraction of turns where a top dialogue state hypothesis was correct (accuracy) or alternatively:
+For quality evaluation one can use either a fraction of turns where a top dialogue state hypothesis was correct (accuracy) or alternatively \cite{DST}:
 \begin{equation}
@@
 
-\subsection{Another update rule}
+\subsection{Another Update Rule}
 Define an update rule:
@@
 
-\section{Dialog optimisation as a reinforcement learn\-ing problem}
-Apart from tracking a state of a dialog, the agent must learn to perform correct actions.
-This can be viewed as a reinforcement learning problem, where the agent is rewarded for performing correct actions and penalised for incorrect ones.
-Agent has a policy (its behaviour) that should be optimized, rewards from the environment, $V$-function (prediction of the future reward given current state $b$) and a model of environment.
+\section{Dialog Optimisation as a Reinforcement Learn\-ing Problem}
+Apart from DST problem, the agent must learn to execute correct actions.
+This can be viewed as a reinforcement learning (RL) problem, where the agent is rewarded for performing correct actions and penalised for incorrect ones.
+Agent has a policy (its behaviour) that should be optimized, rewards from the environment, $V$-function (prediction of the future reward given current state $b$) and a model of environment sometimes.
 
-$Q$-function is also defined to determines how good is it to take action $a$ in belief state $b$ and then follow a policy $\pi$. Having an optimal $Q$-function one can choose an optimal action.
+$Q$-function is also defined to determine how good is it to take action $a$ in belief state $b$ and then follow a policy $\pi$. Having an optimal $Q$-function one can choose an optimal action.
 
@@
 
-The optimal value of Q is:
+The optimal value of Q is \cite{Sutton1998}:
 \begin{align}
@@
 
-\subsection{Allowed actions for execution}
+\subsection{Actions Allowed for Execution}
 
-For optimization one can restrict some actions to be executed as an agent should be not allowed to execute those actions which are clearly implausible.
+For optimization one can restrict some actions to be executed as agent should be not allowed to execute those actions which are clearly implausible.
 For example, after user requests a restaurant's address an agent should not say ``hello”.
@@
 
-\subsection{Monte-Carlo policy}
+\subsection{Monte-Carlo Policy}
 
-Monte-Carlo control algorithm finds an optimal policy for a reinforcement learning problem.
+Monte-Carlo control (MCC) algorithm \cite{Sutton1998} finds an optimal policy for a RL problem.
 
@@
     \centering
-    \includegraphics[width=10cm]{MCC.png}
-    \caption{}
+    \includegraphics[width=12cm]{MCC.png}
+    \caption{Monte Carlo Control algorithm \cite{Sutton1998}}
     \label{fig:MCC}
@@
 \pagebreak
-
 %===========================================================
@@
 
-where $\gamma$ is a discount factor,  $r$ is an immediate reward and $\max_a Q(s',a)$ is the maximum Q-value of the next state. Bellman equation is used to update the Q-function iteratively:
+\noindent where $\gamma$ is a discount factor,  $r$ is an immediate reward and $\max_a Q(s',a)$ is the maximum Q-value of the next state.
+Bellman equation is used to update the Q-function iteratively \cite{DQN}:
 
@@
 
-In comparison to Monte-Carlo this approach is more efficient as it does not keep track of all the past states and remembers only the current one, because only Q-value of the current state is used for the next step update.
-Q-Network trains faster and better then a Monte-Carlo approach (TODO 96\% vs.
-76\% Figures \ref{fig:ex}, \ref{fig:mc}).
+In comparison to MCC this approach is more efficient as it does not keep track of all the past states and remembers only the current one, because only Q-value of the current state is used for the next step update.
+Q-Network trains faster and better than a MCC approach (96\% vs.
+76\% Figure \ref{fig:ex-no-mc}).
 
-Two reasons for benefits of DQN over Monte-Caro are: experience replay and delayed Q-targets.
+Two reasons for benefits of DQN over MCC are: experience replay and delayed Q-targets \cite{DQN}.
 For experience replay one stores tuples ${s, a, r, s^\prime}$ in the buffer, where $s^\prime$ is the following dialog state and samples batches from the it.
@@
 
-Let's train a Q-network without experience replay.
-In this case reward was almost 3 times lower TODO (11.23 and 4.5 Figure \ref{fig:no-ex}, Table \ref{tab:replay-no-replay}).
+Additionally Q-network was trained without experience replay.
+In this case reward was almost 3 times lower -- 11.23 vs.
+4.5 (Figure \ref{fig:ex-no-mc}, Table \ref{tab:replay-no-replay}).
 With increased variance the $Q$-network became more unstable.
@@
      \hline\hline
-     Experience replay & TODO 11.23 +- 0.97 & 93.00 +- 3.56 & 7.37 +- 0.53 \\ 
+     Experience replay & 11.23 +- 0.97 & 93.00 +- 3.56 & 7.37 +- 0.53 \\ 
      \hline
@@
      \hline
+     Monte-Carlo & 7.07 +- 1.48 & 76.00 +- 5.96 & 8.13 +- 0.61 \\
+     \hline
     \end{tabular}
-    \caption{}
+    \caption{Results of DQN training}
     \label{tab:replay-no-replay}
@@
 
-TODO change images
 \begin{figure}[!htb]
-    \minipage{0.32\textwidth}
-      \includegraphics[width=\linewidth]{env2-CamRestaurants-experience.png}
-      \caption{Experience Replay}
-      \label{fig:ex}
-      \endminipage\hfill
-    \minipage{0.32\textwidth}
-      \includegraphics[width=\linewidth]{env2-CamRestaurants-experience.png}
-      \caption{No Experience Replay}
-      \label{fig:no-ex}
-      \endminipage\hfill
-    \minipage{0.32\textwidth}%
-      \includegraphics[width=\linewidth]{env2-CamRestaurants-mc.png}
-      \caption{Monte-Carlo}
-      \label{fig:mc}
-    \endminipage
+    \centering
+    \includegraphics[width=\linewidth]{ex-no-mc.png}
+    \caption{DQN,\quad DQN without Experience Replay,\quad MCC}
+    \label{fig:ex-no-mc}
 \end{figure}
@@
 System sometimes gets caught in the loop and keeps asking the same question.
-The reason might be that semantic decoder didn't work well and haven't recog\-nised an input correctly, so the system continues to request more information.
-TODO get example
+The reason might be malfunction of a semantic decoder with wrong inpure recognition and therefore system continues to request more information.
 
-There was no instances in the dataset where the user wouldn't care which price range is there, and the dialog is ruined when this case is encountered (Dialog \ref{diag:with-price} vs Dialog \ref{diag:without-price}).
+\begin{figure}[!htb]
+    \renewcommand\figurename{Dialog}
+    \centering
+    \includegraphics[width=10cm]{infinite\ loop.png}
+    \caption{Infinite Loop in the dialog}
+    \label{diag:loop}
+\end{figure}
+
+There were no instances in the dataset where user wouldn't care about the price range, and the dialog gets ruined when such case is encountered (Dialog \ref{diag:with-price} vs Dialog \ref{diag:without-price}).
 This might happen due to insufficient semantic decoder knowledge and not enough of such samples in the database.
@@
 \begin{figure}[!htb]
+    \renewcommand\figurename{Dialog}
     \minipage{0.49\textwidth}
@@
 
-\begin{figure}
+\begin{figure}[!htb]
+    \renewcommand\figurename{Dialog}
     \centering
-    \includegraphics[width=5cm]{chinese\ negation.png}
+    \includegraphics[width=10cm]{chinese\ negation.png}
     \caption{}
@@
 \pagebreak
-
 \section{Information Gain}
@@
 One can also substitute overall reward $r$ with $r = r + \nu r_i$, here $r_i$ is an additional intrinsic reward, and $\nu$ is a hyperparameter.
-This reward is called information gain and it measures how much additional information the system gets about user needs after one turn.
-Let action be of form $\{request, confirm,$ $select\}\_\xi$, $p_{\xi}$ and ${p_{\xi}}^\prime$ --- belief distributions over possible values of $\xi$ in current and next belief state.
-Informa\-tion gain then is the distance between these distributions measured by with Jensen-Shannon di\-vergence distance:
+This reward is called information gain (IG) and it measures how much additional information the system gets about user needs after one turn.
+Let action be of form $\{request,$  $confirm,$ $select\}\_\xi$, $p_{\xi}$ and ${p_{\xi}}^\prime$ --- belief distributions over possible values of $\xi$ in current and next belief state.
+IG is the distance between these distributions measured by with Jensen-Shannon di\-vergence distance:
 
@@
 
-When distance is big enough (higher than a threshold) we add infor\-mation gain to the reward, but only when the action was of the form $\{request,$ $ confirm, select\}$.
-Intuitively it also doesn't make sense to calculate infor\-mation gain for actions like $bye, inform$ as they do not provide any additonal infor\-mation on user needs and therefore do not lead to the accomplishment of the dialog goal.
+When distance is big enough (higher than a threshold) we add IG to the reward, but only when the action was of the form $\{request,$ $ confirm, select\}$.
+Intuitively it also doesn't make sense to calculate IG for actions like $bye,$ $inform$ as they do not provide any additonal infor\-mation on user needs and therefore do not lead to the accomplishment of the dialog goal.
 It is beneficial to use intrinsic rewards in addition as they encourage system to get more useful information from users by requesting more from them, selecting the most relevant one information and confirming the choices.
 
-Here one can see the difference in training with information gain and without it.
-Information gain improves the speed of training and slightly improves an overall performance from $96\%$  to $98.4\%$ (see Figure 7, 8).
+Here one can see the difference in training with IG and without it. IG improves the speed of training and slightly improves an overall performance from $96\%$  to $98.4\%$ (see Figure 7, 8).
 
@@
       \includegraphics[width=\linewidth]{env3-ig-CamRestaurants.png}
-      \caption{Information Gain}
+      \caption{Information gain}
     \endminipage\hfill
@@
       \includegraphics[width=\linewidth]{env3-CamRestaurants.png}
-      \caption{No Information Gain}
+      \caption{No information gain}
     \endminipage
@@
 One can also substitute $r = r + \nu r_i$.
-However one should be cautious --- if the value $\mu$ is too high, we push the sytem to gather more and more information and discourage it from making a choice and informing a user about it.
+However it should be done cautious --- if the value $\mu$ is too high, the sytem is pushed to gather more and more information and is discouraged from making a choice and informing a user about it.
 
@@
 This report gives a brief introduction to the construction of spoken dialog systems.
-It contains research on different approaches and algorithms used for end-to-end spoken dialog system, that are build via modular approach.
-Such system was built with the use of different deep learning approaches, all approaches for each of the modules were studied and compared, several dialogs were conducted with the system.
+It contains research on different approaches and algorithms used for systems, that are build via modular approach.
+One such system was built with the use of different deep learning approaches, all approaches for each of the modules were studied and compared, several dialogs were conducted with the system.
 The results of the study were described and presented.
